---
title: "DRAFT: Internal MOSAIC Sensemaking"
subtitle: "Do Not Share"
author: "Catalyst California"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 3
    toc_float: yes
date: "`r format(Sys.time(), '%d %B %Y')`"
---


```{css, echo=FALSE}
.highcharts-container,
.highcharts-root {
  overflow: visible !important;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

### Set up workspace ####
#install.packages("extrafont")
library(extrafont)
library(here)
library(dplyr)
library(RPostgres)
library(formattable)
library(svglite)
library(stringr)
library(tidyr)
library(showtext)
library(scales)
library(highcharter)

options(scipen=999)

# set global options to ensure that comma separator is a comma for highchart graphs

lang <- getOption("highcharter.lang")
lang$thousandsSep <- ","
options(highcharter.lang = lang)

# load the PostgreSQL driver
source("W:\\RDA Team\\R\\credentials_source.R")
# load fx
source(here("MOSAIC/Functions/viz_fx.R"))
con <- connect_to_db("mosaic")
con2 <- connect_to_db("racecounts")



# set variables
rc_schema <- 'v7'
curr_yr <- '2025'

####################
#### Style Guide ###
####################

## COLORS: There is only blue ##
black <- "#000000"
mainblue <- "#22BCB8"
textgrey <- "#919191"
linecolor<-"#FF0000"

# define fonts in chart
font_title <- "Arial"
font_subtitle <- "Arial"
font_caption <- "Arial"
font_bar_label <- "Arial"
font_axis_label <- "Arial"
font_table_text<-"Arial"

####################################
########Highchart theme#############
####################################

#### Define Highcharts theme #### 
tnp_theme <- hc_theme(
  colors = c("#22BCB8"  # main blue
            
  ),
  chart = list(
    backgroundColor = "white",
    style = list(
      fontFamily = font_subtitle
    ) 
  ),
  plotOptions = 
    list(
      line = 
        list(
          marker = 
            list(
              enabled = FALSE
            )
        )
    ),
  title = list(
    style = list(
      color = black,
      fontFamily = font_title,
      textAlign="left",
      fontsize='200px'
    )
  ),
  subtitle = list(
    style = list(
      color = black, # medium grey
      fontFamily = font_subtitle,
      textAlign="left",
      fontsize='100px'
    )
  ),
  caption = list(
    style = list(
      color = black,
      fontFamily = font_caption,
      textAlign = "left",
      fontsize="12px"
    ),
    useHTML = TRUE
  ),
  axis = list(
    style = list(
      color = black,
      fontFamily = font_axis_label,
      fontSize='50px'
    )
  ),
  legend = list(
    itemStyle = list(
      fontFamily = font_axis_label,
      color = black
    ),
    itemHoverStyle = list(
      fontFamily = font_table_text,
      color = black
    )
  )
)

### Info: Race 
lang <- getOption("highcharter.lang")
lang$thousandsSep <- ","
options(highcharter.lang = lang)

```

```{r metadata, include=FALSE}

### Metadata: Dummy is v7 county/state, WILL NEED TO BE REPLACED WITH MOSAIC INDICATOR METADATA TABLE
metadata <- dbGetQuery(con2, paste0("SELECT * FROM ", rc_schema, ".", "arei_indicator_list_cntyst"))
# metadata$api_name

acs_racenote <- "Data Note: Subgroups are Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities."

pums_racenote <- "Data Note: Subgroups are Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities. Subgroups are defined by Ancestry."

cdc_racenote <- "Data Note: Subgroups include foreign-born birthing parents only and are defined by country of birth."

hmda_racenote <- "Data Note: Subgroups Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities. This data differs from regular RC because it is based on only applicant race, not a combination of applicant and co-applicant."


```

# Introduction {#intro}
<span style="font-size: 15px;">These are draft visuals for internal sensemaking only. These visuals do not reflect what final visuals on MOSAIC report page will look or behave like.</style>
<br><br>
<span style="font-size: 15px;">Note that the Asian/NHPI/Total rates displayed as vertical lines are sourced from our 'regular' RC data for the most part. The data sources and methods vary slightly by necessity from those of MOSAIC analysis. Once we select indicators, we will recalculate the Total and raced rates using the MOSAIC sources/methods for more direct comparison. Some charts may have Asian, NHPI, and other race bars - those reflect MOSAIC calculations of those rates.</style>
<br><br>
<span style="font-size: 15px;">The data sources listed here are mostly correct, the actual data years for the MOSAIC analysis may vary based on data availability. Some datasets are older than shown here (ACS), some are more updated (CHIS). All data years will be accurately displayed on report page.</style>


# Asian Disaggregation {#asian}

```{r asian-charts, include=FALSE}

#### INTERNET ####
# # Run metadata fx - run metadata$api_name if need to find indicator variable name
# internet_meta <- metadata_fx("econ", "internet", "asian", metadata)
# 
# # Run data fx
# internet_list <- data_fx(internet_meta, "asian")
# 
# # Run chart fx
# internet_plot <- chart_fx(internet_list, internet_meta, "asian", acs_racenote)


# #### CONNECTED YOUTH ####
# # Run metadata fx
# youth_meta <- metadata_fx("econ", "connected_youth", "asian", metadata, "2025")
# 
# # Run data fx
# youth_list <- data_fx(youth_meta, "asian", "v7")
# 
# # Run chart fx
# youth_plot <- chart_fx(youth_list, youth_meta, "asian", acs_racenote)


#### OFFICIALS ####
# Run metadata fx
officials_meta <- metadata_fx("econ", "officials", "asian", metadata, "2025")

# Run data fx
officials_list <- data_fx(officials_meta, "asian", "v7")

# Run chart fx
officials_plot <- chart_fx(officials_list, officials_meta, "asian", pums_racenote)


#### OVERCROWDED ####
# Run metadata fx
overcrowd_meta <- metadata_fx("hous", "overcrowded", "asian", metadata, "2023")

# Run data fx
overcrowd_list <- data_fx(overcrowd_meta, "asian", "v5")

# Run chart fx
overcrowd_plot <- chart_fx(overcrowd_list, overcrowd_meta, "asian", acs_racenote)


#### HOMEOWNERSHIP ####
# Run metadata fx
homeown_meta <- metadata_fx("hous", "homeownership", "asian", metadata, "2023")

# Run data fx
homeown_list <- data_fx(homeown_meta, "asian", "v5")

# Run chart fx
homeown_plot <- chart_fx(homeown_list, homeown_meta, "asian", acs_racenote)


#### RENTER COST BURDEN ####
# Run metadata fx
renter_meta <- metadata_fx("hous", "cost_burden_renter", "asian", metadata, "2025")

# Run data fx
renter_list <- data_fx(renter_meta, "asian", "v7")

# Run chart fx
renter_plot <- chart_fx(renter_list, renter_meta, "asian", acs_racenote)


#### OWNER COST BURDEN ####
# Run metadata fx
owner_meta <- metadata_fx("hous", "cost_burden_owner", "asian", metadata, "2025")

# Run data fx
owner_list <- data_fx(owner_meta, "asian", "v7")

# Run chart fx
owner_plot <- chart_fx(owner_list, owner_meta, "asian", acs_racenote)


#### UNINSURED ####
# Run metadata fx
uninsured_meta <- metadata_fx("hlth", "health_insurance", "asian", metadata, "2023")

# Run data fx
uninsured_list <- data_fx(uninsured_meta, "asian", "v5")

# Run chart fx
uninsured_plot <- chart_fx(uninsured_list, uninsured_meta, "asian", cdc_racenote)


#### LOW BIRTHWEIGHT ####
# Run metadata fx
lbw_meta <- metadata_fx("hlth", "low_birthweight", "asian", metadata, "2025")

# Run data fx
lbw_list <- data_fx(lbw_meta, "asian", "v7")

# Run chart fx
lbw_plot <- chart_fx(lbw_list, lbw_meta, "asian", cdc_racenote)


#### VOTER ENGAGEMENT ####  Does not work with functions bc not in RC v7
# Run metadata fx
voter_data <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".", "asian_demo_voter_engagement_state_2025")) %>%
    pivot_longer(
      cols = -state_id,
      names_to = c("subgroup", ".value"),
      names_pattern = "^(.+)_(raw|rate)$"
    ) %>%
    # make clean subgroup labels
    mutate(subgroup_label = gsub("_aoic", "", subgroup)) %>%
    mutate(subgroup_label = str_to_title(gsub("_", " ", subgroup_label))) %>%
    na.omit()

race_ <- 'asian'
raw_descriptor <- 'US citizens always or frequently engaged'
asbest <- 'max'
title_text <- 'Voter Engagement'
subtitle_text <- 'US citizens always or frequently engaged in national, state, and local elections'
datasource <- 'CHIS 2019-2024'
racenote <- "Data Note: Subgroups are Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities."


# pulled from: W:\Project\RACE COUNTS\2025_v7\MOSAIC\QA_Sheet_Sensemaking.docx
race_rate <- 54.6
total_rate <- 65.1
  
  # generate race rate label
  race_label <- case_when(
    race_ %in% c('asian','Asian') ~ "Asian",
    TRUE ~ "Native Hawaiian or Pacific Islander"
  )
  
  # dynamic height based on number of bars bc there are many more Asian subgroups
  chart_height <- case_when(
    nrow(voter_data) >= 13 ~ 700,
    nrow(voter_data) >= 8  ~ 500,
    TRUE                     ~ 300
  )
  
  # build chart
voter_plot <- b_chart <- hchart(
    voter_data %>% arrange(if (asbest == 'max') rate else -rate),   # sort asc/desc depending on asbest value
    type = "bar", 
    hcaes(y = round(rate, 1), x = subgroup_label), 
    color = mainblue,
    tooltip = list(   # Add % when max value <= 100 (aka values are percents)
      pointFormat = paste0(
        "Rate: {point.rate:.1f}", if (max(voter_data$rate, na.rm = TRUE) > 100) "" else "%",
        "<br>", str_to_sentence(raw_descriptor), ": {point.raw:,.0f}"
      )
    ),
  ) %>% 
    
    hc_tooltip(crosshairs = TRUE) %>% 
    
    hc_xAxis(title = list(text = ""),
             labels=list(style=list(fontSize='15px'))
    ) %>% 
    
    hc_yAxis(title = list(text = ""),
             labels = list(  # # Add % when max value <= 100 (aka values are percents)
               format = if (max(race_rate, na.rm = TRUE) > 100) "{value}" else "{value}%"
             ), list(step = 1), padding = 0,    # this is for rate indicators only adding % signs
             max = min(1000, max(voter_data$rate, na.rm = TRUE) * 1.1),        # data max + 10% buffer, capped at 100
             plotLines = list(
               list(
                 value = total_rate,   # total value
                 color = "red",
                 width = 2,
                 dashStyle = "Solid",     # "Solid", "Dash", "Dot", "DashDot"
                 zIndex = 5,              # draw on top of bars
                 label = list(
                   text = paste0(
                     "Total Rate: ",  # Add % when max value <= 100 (aka values are percents)
                     round(total_rate, 1),
                     if (max(total_rate, na.rm = TRUE) > 100) "" else "%"
                   ),
                   align = "right",       # place label to the left of the line
                   verticalAlign = "bottom",
                   rotation = 0,          # make labels horizontal
                   y = -15,
                   style = list(color = "red", fontWeight = "bold")
                 )
               ),
               list(
                 value = race_rate,   # race group value
                 color = "black",
                 width = 2,
                 dashStyle = "Solid",     # "Solid", "Dash", "Dot", "DashDot"
                 zIndex = 5,              # draw on top of bars
                 label = list(
                   text = paste0(
                     race_label, " Rate: ",   # Add % when max value <= 100 (aka values are percents)
                     round(race_rate, 1),
                     if (max(race_rate, na.rm = TRUE) > 100) "" else "%"
                   ),
                   align = "right",       # place label to the left of the line
                   verticalAlign = "bottom",
                   rotation = 0,          # make labels horizontal
                   y = 5,
                   style = list(color = "black", fontWeight = "bold")
                 )
               )  
             )
    ) %>%
    
    hc_plotOptions(
      bar = list(
        groupPadding = 0.05,   # space between groups (default 0.2)
        pointPadding = 0.01    # width of bars
      )
    ) %>%
    
    # title elements
    hc_title(
      text = paste0(title_text),
      #margin = 20,
      align = "left", 
      style = list(useHTML = TRUE, fontWeight = "bold", fontSize='30px')
    ) %>% 
    
    hc_subtitle(
      text = paste0(subtitle_text),
      align = "left", y = 50
    ) %>% 
    
    hc_caption(
      text = paste0("<br>Data Source: ", datasource,
                    "<br>",racenote),
      style=list(fontSize='15px')
      #align = "left"
    ) %>% 
    
    hc_size(height = chart_height) %>%
    
    hc_add_theme(tnp_theme)


```

### Officials & Managers
```{r, echo=FALSE}
officials_plot
```

### Overcrowded Housing
```{r, echo=FALSE}
overcrowd_plot
```

### Homeownership
```{r, echo=FALSE}
homeown_plot
```

### Renter Cost Burden
```{r, echo=FALSE}
renter_plot
```

### Owner Cost Burden
```{r, echo=FALSE}
owner_plot
```

### Uninsured
```{r, echo=FALSE}
uninsured_plot
```

### Voter Engagement
```{r, echo=FALSE}
voter_plot
```


# NHPI Disaggregation {#nhpi}

```{r nhpi-charts, include=FALSE}

#### INTERNET ####
# # Run metadata fx - run metadata$api_name if need to find indicator variable name
# internet_meta <- metadata_fx("econ", "internet", "nhpi", metadata)
# 
# # Run data fx
# internet_list <- data_fx(internet_meta, "nhpi")
# 
# # Run chart fx
# internet_plot <- chart_fx(internet_list, internet_meta, "nhpi", acs_racenote)


#### CONNECTED YOUTH ####
# Run metadata fx
youth_meta <- metadata_fx("econ", "connected_youth", "nhpi", metadata, "2025")

# Run data fx
youth_list <- data_fx(youth_meta, "nhpi", "v7")

# Run chart fx
youth_plot <- chart_fx(youth_list, youth_meta, "nhpi", acs_racenote)


#### OVERCROWDED ####
# Run metadata fx
overcrowd_meta <- metadata_fx("hous", "overcrowded", "nhpi", metadata, "2023")

# Run data fx
overcrowd_list <- data_fx(overcrowd_meta, "nhpi", "v5")

# Run chart fx
overcrowd_plot <- chart_fx(overcrowd_list, overcrowd_meta, "nhpi", acs_racenote)


#### HOMEOWNERSHIP ####
# Run metadata fx
homeown_meta <- metadata_fx("hous", "homeownership", "nhpi", metadata, "2023")

# Run data fx
homeown_list <- data_fx(homeown_meta, "nhpi", "v5")

# Run chart fx
homeown_plot <- chart_fx(homeown_list, homeown_meta, "nhpi", acs_racenote)


#### DENIED MORTGAGE ####
# Run metadata fx
mtg_meta <- metadata_fx("hous", "denied_mortgages", "nhpi", metadata, "2025")

# Run data fx
mtg_list <- data_fx(mtg_meta, "nhpi", "v7")

# Run chart fx
mtg_plot <- chart_fx(mtg_list, mtg_meta, "nhpi", hmda_racenote)


### UNINSURED ####
# Run metadata fx
uninsured_meta <- metadata_fx("hlth", "health_insurance", "nhpi", metadata, "2023")

# Run data fx
uninsured_list <- data_fx(uninsured_meta, "nhpi", "v5")

# Run chart fx
uninsured_plot <- chart_fx(uninsured_list, uninsured_meta, "nhpi", acs_racenote)


#### LOW BIRTHWEIGHT ####
# Run metadata fx
lbw_meta <- metadata_fx("hlth", "low_birthweight", "nhpi", metadata, "2025")

# Run data fx
lbw_list <- data_fx(lbw_meta, "nhpi", "v7")

# Run chart fx
lbw_plot <- chart_fx(lbw_list, lbw_meta, "nhpi", cdc_racenote)


```

### Connected Youth
```{r, echo=FALSE}
youth_plot
```

### Uninsured
```{r, echo=FALSE}
uninsured_plot
```

### Low Birthweight Births
```{r, echo=FALSE}
lbw_plot
```

### Overcrowded Housing
```{r, echo=FALSE}
overcrowd_plot
```

### Homeownership
```{r, echo=FALSE}
homeown_plot
```

### Denied Mortgages
```{r, echo=FALSE}
mtg_plot
```
