---
title: "DRAFT: Internal MOSAIC Sensemaking"
subtitle: "Do Not Share"
author: "Catalyst California"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 3
    toc_float: yes
date: "`r format(Sys.time(), '%d %B %Y')`"
---

<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

<style>

/* Set TOC formatting for active items */
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover{
  color: #3A207D !important;     /* meteorite */
  background-color: white;
  font-weight: 600;
  font-size: 23;
}

/* set addt'l unordered (bulleted) list formatting (override formatting set in css) */
.large-list ul {
  font-family: Inter, sans-serif;
  color: meteorite;
  font-size: 16px;
}

/* set accordion label formatting */
summary {
  color: #3A207D !important;    /* Change the color to meteorite */
}

</style>


```{css, echo=FALSE}
.highcharts-container,
.highcharts-root {
  overflow: visible !important;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

### Set up workspace ####
#install.packages("extrafont")
library(extrafont)
library(here)
library(dplyr)
library(RPostgres)
library(formattable)
library(svglite)
library(stringr)
library(tidyr)
library(showtext)
library(scales)
library(highcharter)

options(scipen=999)

# set global options to ensure that comma separator is a comma for highchart graphs

lang <- getOption("highcharter.lang")
lang$thousandsSep <- ","
options(highcharter.lang = lang)

# load the PostgreSQL driver
source("W:\\RDA Team\\R\\credentials_source.R")
# load fx
source(here("MOSAIC/Functions/viz_fx.R"))
con <- connect_to_db("mosaic")
con2 <- connect_to_db("racecounts")



# set variables
rc_schema <- 'v7'
curr_yr <- '2025'

####################
#### Style Guide ###
####################

## COLORS: There is only blue ##
black <- "#000000"
mainblue <- "#22BCB8"
textgrey <- "#919191"
linecolor<-"#FF0000"

# define fonts in chart
font_title <- "Arial"
font_subtitle <- "Arial"
font_caption <- "Arial"
font_bar_label <- "Arial"
font_axis_label <- "Arial"
font_table_text<-"Arial"

####################################
########Highchart theme#############
####################################

#### Define Highcharts theme #### 
tnp_theme <- hc_theme(
  colors = c("#22BCB8"  # main blue
            
  ),
  chart = list(
    backgroundColor = "white",
    style = list(
      fontFamily = font_subtitle
    ) 
  ),
  plotOptions = 
    list(
      line = 
        list(
          marker = 
            list(
              enabled = FALSE
            )
        )
    ),
  title = list(
    style = list(
      color = black,
      fontFamily = font_title,
      textAlign="left",
      fontsize='200px'
    )
  ),
  subtitle = list(
    style = list(
      color = black, # medium grey
      fontFamily = font_subtitle,
      textAlign="left",
      fontsize='100px'
    )
  ),
  caption = list(
    style = list(
      color = black,
      fontFamily = font_caption,
      textAlign = "left",
      fontsize="12px"
    ),
    useHTML = TRUE
  ),
  axis = list(
    style = list(
      color = black,
      fontFamily = font_axis_label,
      fontSize='50px'
    )
  ),
  legend = list(
    itemStyle = list(
      fontFamily = font_axis_label,
      color = black
    ),
    itemHoverStyle = list(
      fontFamily = font_table_text,
      color = black
    )
  )
)

### Info: Race 
lang <- getOption("highcharter.lang")
lang$thousandsSep <- ","
options(highcharter.lang = lang)

```

```{r import-metadata, include=FALSE}

### Metadata: Dummy is v7 county/state, WILL NEED TO BE REPLACED WITH MOSAIC INDICATOR METADATA TABLE
metadata <- dbGetQuery(con2, paste0("SELECT * FROM ", rc_schema, ".", "arei_indicator_list_cntyst"))

# metadata$api_name

```

```{r asian-charts, include=FALSE}

acs_racenote <- "Data Note: Subgroups are Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities."

pums_racenote <- "Data Note: Subgroups are Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities. Subgroups are defined by Ancestry."

cdc_racenote <- "Data Note: Subgroups include foreign-born birthing parents only and are defined by country of birth."


#### INTERNET ####
# # Run metadata fx - run metadata$api_name if need to find indicator variable name
# internet_meta <- metadata_fx("econ", "internet", "asian", metadata)
# 
# # Run data fx
# internet_list <- data_fx(internet_meta, "asian")
# 
# # Run chart fx
# internet_plot <- chart_fx(internet_list, internet_meta, "asian", acs_racenote)


#### OVERCROWDED ####
# Run metadata fx
overcrowd_meta <- metadata_fx("hous", "overcrowded", "asian", metadata, "2023")

# Run data fx
overcrowd_list <- data_fx(overcrowd_meta, "asian", "v5")

# Run chart fx
overcrowd_plot <- chart_fx(overcrowd_list, overcrowd_meta, "asian", acs_racenote)


#### HOMEOWNERSHIP ####
# Run metadata fx
homeown_meta <- metadata_fx("hous", "homeownership", "asian", metadata, "2023")

# Run data fx
homeown_list <- data_fx(homeown_meta, "asian", "v5")

# Run chart fx
homeown_plot <- chart_fx(homeown_list, homeown_meta, "asian", acs_racenote)


#### RENTER COST BURDEN ####
# Run metadata fx
renter_meta <- metadata_fx("hous", "cost_burden_renter", "asian", metadata, "2025")

# Run data fx
renter_list <- data_fx(renter_meta, "asian", "v7")

# Run chart fx
renter_plot <- chart_fx(renter_list, renter_meta, "asian", acs_racenote)


#### OWNER COST BURDEN ####
# Run metadata fx
owner_meta <- metadata_fx("hous", "cost_burden_owner", "asian", metadata, "2025")

# Run data fx
owner_list <- data_fx(owner_meta, "asian", "v7")

# Run chart fx
owner_plot <- chart_fx(owner_list, owner_meta, "asian", acs_racenote)


#### UNINSURED ####
# Run metadata fx
uninsured_meta <- metadata_fx("hlth", "health_insurance", "asian", metadata, "2023")

# Run data fx
uninsured_list <- data_fx(uninsured_meta, "asian", "v5")

# Run chart fx
uninsured_plot <- chart_fx(uninsured_list, uninsured_meta, "asian", cdc_racenote)


#### LOW BIRTHWEIGHT ####
# Run metadata fx
lbw_meta <- metadata_fx("hlth", "low_birthweight", "asian", metadata, "2025")

# Run data fx
lbw_list <- data_fx(lbw_meta, "asian", "v7")

# Run chart fx
lbw_plot <- chart_fx(lbw_list, lbw_meta, "asian", cdc_racenote)


#### VOTER ENGAGEMENT ####  Does not work with functions bc not in RC v7
# Run metadata fx
#voter_meta <- metadata_fx("demo", "voter_engagement", "asian", metadata)

# Run data fx
#voter_list <- data_fx(voter_meta, "asian")

# Run chart fx
#voter_plot <- chart_fx(voter_list, voter_meta, "asian", cdc_racenote)


```

# Introduction {#intro}
#### These are draft visuals for internal sensemaking only. These visuals do not reflect what final visuals on MOSAIC report page will look or behave like.

#### Note that the Asian/NHPI/Total rates displayed as vertical lines are sourced from our 'regular' RC data. The data sources and methods vary slightly by necessity from those of MOSAIC analysis. Once we select indicators, we will recalculate the Asian/NHPI/Total rates using the MOSAIC sources/methods for more direct comparison.

#### The data sources listed here are correct, the actual data years for the MOSAIC analysis may vary based on data availability. Some datasets are older than shown here (ACS), some are more updated (CHIS). All data years will be accurately displayed on report page.


# Asian Disaggregation {#asian}

### Overcrowded Housing
```{r, echo=FALSE}
overcrowd_plot
```

### Homeownership
```{r, echo=FALSE}
homeown_plot
```

### Renter Cost Burden
```{r, echo=FALSE}
renter_plot
```

### Owner Cost Burden
```{r, echo=FALSE}
owner_plot
```

### Uninsured
```{r, echo=FALSE}
uninsured_plot
```



```{r nhpi-charts, include=FALSE}

acs_racenote <- "Data Note: Subgroups are Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities."

pums_racenote <- "Data Note: Subgroups are Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities. Subgroups are defined by Ancestry."

cdc_racenote <- "Data Note: Subgroups include foreign-born birthing parents only and are defined by country of birth."

hmda_racenote <- "Data Note: Subgroups Latinx-inclusive and Alone or in Combination, which means they include all people who identify as the subgroup\nincluding in combination with other subroups, races, and/or ethnicities."

#### INTERNET ####
# # Run metadata fx - run metadata$api_name if need to find indicator variable name
# internet_meta <- metadata_fx("econ", "internet", "nhpi", metadata)
# 
# # Run data fx
# internet_list <- data_fx(internet_meta, "nhpi")
# 
# # Run chart fx
# internet_plot <- chart_fx(internet_list, internet_meta, "nhpi", acs_racenote)

#### CONNECTED YOUTH ####
# Run metadata fx
youth_meta <- metadata_fx("econ", "connected_youth", "nhpi", metadata, "2025")

# Run data fx
youth_list <- data_fx(youth_meta, "nhpi", "v7")

# Run chart fx
youth_plot <- chart_fx(youth_list, youth_meta, "nhpi", acs_racenote)


#### UNINSURED ####
# # Run metadata fx 
# uninsured_meta <- metadata_fx("hlth", "health_insurance", "nhpi", metadata)
# 
# # Run data fx
# uninsured_list <- data_fx(uninsured_meta, "nhpi")
# 
# # Run chart fx
# uninsured_plot <- chart_fx(uninsured_list, uninsured_meta, "nhpi", cdc_racenote)


#### OVERCROWDED ####
# Run metadata fx
overcrowd_meta <- metadata_fx("hous", "overcrowded", "nhpi", metadata, "2023")

# Run data fx
overcrowd_list <- data_fx(overcrowd_meta, "nhpi", "v5")

# Run chart fx
overcrowd_plot <- chart_fx(overcrowd_list, overcrowd_meta, "nhpi", acs_racenote)


#### HOMEOWNERSHIP ####
# Run metadata fx
homeown_meta <- metadata_fx("hous", "homeownership", "nhpi", metadata, "2023")

# Run data fx
homeown_list <- data_fx(homeown_meta, "nhpi", "v5")

# Run chart fx
homeown_plot <- chart_fx(homeown_list, homeown_meta, "nhpi", acs_racenote)


#### DENIED MORTGAGE ####
# Run metadata fx
mtg_meta <- metadata_fx("hous", "denied_mortgages", "nhpi", metadata, "2025")

# Run data fx
mtg_list <- data_fx(mtg_meta, "nhpi", "v7")

# Run chart fx
mtg_plot <- chart_fx(mtg_list, mtg_meta, "nhpi", hmda_racenote)


#### LOW BIRTHWEIGHT ####
# Run metadata fx
lbw_meta <- metadata_fx("hlth", "low_birthweight", "nhpi", metadata, "2025")

# Run data fx
lbw_list <- data_fx(lbw_meta, "nhpi", "v7")

# Run chart fx
lbw_plot <- chart_fx(lbw_list, lbw_meta, "nhpi", cdc_racenote)


```

# NHPI Disaggregation {#nhpi}

### Connected Youth
```{r, echo=FALSE}
youth_plot
```

### Low Birthweight Births
```{r, echo=FALSE}
lbw_plot
```

### Overcrowded Housing
```{r, echo=FALSE}
overcrowd_plot
```

### Homeownership
```{r, echo=FALSE}
homeown_plot
```

### Denied Mortgages
```{r, echo=FALSE}
mtg_plot
```
