---
title: ""
output: 
  html_document:
    self_contained: true
---
<style>
.html-widget {
    width: 100% !important;
    height: 975px !important;
    max-width: 100% !important;
}

body { 
  margin: 0; 
}

/* Hide default Highcharts drillup button */
.highcharts-drillup-button {
    display: none !important;
}

</style>

<!-- Add this SECOND style block after the first -->
<style>
html, body {
    padding: 0 !important;
    margin: 0 !important;
    overflow-x: hidden !important;
}
.container-fluid,
.main-container,
.row {
    padding-left: 0 !important;
    padding-right: 0 !important;
    margin-left: 0 !important;
    margin-right: 0 !important;
    max-width: 100% !important;
}
</style>

```{r, include=FALSE}
# load libraries and test data
library(here)
library(stringr)

source("W:\\RDA Team\\R\\credentials_source.R")
source(here("MOSAIC", "charts", "chart_style_vars.R"))
con <- connect_to_db("mosaic")

knitr::opts_chunk$set(echo = FALSE)

```

```{r}
##### DOCUMENTATION (can skip to next section for indicator inputs) #####
### Requirements for chart behavior: drilldown_js
# total_rate - single variable to establish vertical plotline; total/state rate  of the indicator  

### Requirements for chart behavior: drillup_js
# none

### Requirements: Chart Metadata
# data_source - e.g., "US Department of Health and Human Services, Centers for Disease Control and Prevention (CDC), National Center for Health Statistics, Division of Vital Statistics, CDC WONDER Online Database (2016-2023)"

# data_description - e.g., "The number of infants born weighing less than 2,500 grams or about 5 lbs 8 oz per 100 live births by race/ethnicity of birthing parent. Estimates for racial groups with fewer than ten births or areas with fewer than 100,000 people were excluded to protect privacy. SWANA data includes only foreign-born birthing parents."

# exception: methodology_link is controlled by setting a global variable in chart_style_vars
# methodology_link - e.g., "<a href='https://catalystcalifornia.github.io/RaceCounts/Methodology/Indicator_Methodology_CountyState.html' target='_blank'>Full Methodology</a>"

# chart_title - e.g., "Low Birthweight by Race/Ethnicity"

# indicator_asbest - e.g., "min" or "max"

### Requirements: Chart data / Dataframe structure
# The ideal shape of the indicator table is long - with columns for group, subgroup, rate, count. It should have both the mosaic group (asian and/or nhpi and total which should include state (subgroup = total) and the RC race subgroups). It should also have Asian/NHPI as a subgroup under total group
# Ideally rates are already rounded in this table or the metadata table includes this (or we make all round to 1 decimal place)
# Ideally it or an accompanying metadata table includes the asbest value for the indicator.
# additional metadata fields include round rate unless we want all indicators to round to one place
# group - name of column with aggregate group name (e.g., Asian, NHPI, Total) - should be all caps
# subgroup - name of column with subgroup names (Chinese, Indian, etc.) - should be all caps
# rate - name of column with indicator raced rates
# raw - name of column with indicator raced counts
```

```{r}
##### Prep: Get data (must change for each indicator) #####
# Get indicator df
indicator_df <- dbGetQuery(con, 
                           sprintf("SELECT * FROM v7.%s_summary_for_charts%s WHERE topic = '%s'",
                                   race_grp,
                                   geolevel_table_suffix,
                                   ind))

dbDisconnect(conn = con)
# Set metadata - see documentation above for examples
indicator_name <- indicator_df$topic[1]
indicator_asbest <- indicator_metadata %>% filter(api_name == indicator_name) %>% select(arei_best) %>% pull()
data_source <- indicator_metadata %>% filter(api_name == indicator_name) %>% select(data_source) %>% pull()
data_description <- indicator_metadata %>% filter(api_name == indicator_name) %>% select(methodology) %>% pull()
chart_title <- indicator_metadata %>% filter(api_name == indicator_name) %>% select(bar_chart_header) %>% pull()
rate_rounding <- indicator_metadata %>% filter(api_name == indicator_name) %>% select(rate_rounding) %>% pull() %>% as.integer()
raw_rounding <- indicator_metadata %>% filter(api_name == indicator_name) %>% select(raw_rounding) %>% pull() %>% as.integer()

# Set total_rate
total_rate <- indicator_df %>% 
  filter(group=="total" & subgroup == "total") %>% 
  select(rate) %>% pull() %>% round(., rate_rounding)

# # need to filter out NA rates/raw and total rate
indicator_df <- indicator_df %>%
  drop_na(rate) 

# Don't change - pulled in from chart_style_vars
methodology_link <- sprintf("<a href='%s' target='_blank'>Full Methodology</a>", mosaic_methodology_link)
```


```{r}
##### Prep: Edit data for chart (don't change) #####

# use indicator best direction to arrange "shared struggle" groups at top
if(indicator_asbest=="min") {
  indicator_df <- indicator_df %>%
    arrange(desc(rate)) %>%
    mutate(rate = round(rate, rate_rounding),
           raw = round(raw, raw_rounding))
  
} else if (indicator_asbest=="max"){
  indicator_df <- indicator_df %>%
    arrange(rate) %>%
    mutate(rate = round(rate, rate_rounding),
           raw = round(raw, raw_rounding))
}

# define df initial chart of drilldown (group and total bars only)
base_df <- indicator_df %>%
  filter(type=="total" | type=="group") %>%
  select(-type) %>%
  mutate(
    color = case_when(
      group == "asian" ~ base_colors[["asian"]],
      group == "pacisl" ~ base_colors[["pacisl"]],
      group == "total" ~ base_colors[["total"]],
      TRUE ~ base_colors[["default"]]))

# reorg indicator df to get asian and pacisl to appear under total drilldown
indicator_df <- indicator_df %>%
  mutate(
    group = case_when(
      group == "pacisl" ~ "other major race groups",
      group == "asian" ~ "other major race groups",
      .default = group
    )
  )

drilldown_df <- indicator_df %>%
  filter(type!="total") %>%
  group_nest(group) %>%
  mutate(
    id = case_when(
      group == "asian subgroups" ~ "asian",
      group == "pacisl subgroups" ~ "pacisl",
      group == "other major race groups" ~ "total",
      .default = group),
    type = "bar",
    color = case_when(
      id == "asian" ~ base_colors[["asian"]],
      id == "total" ~ base_colors[["total"]],
      id == "pacisl" ~ base_colors[["pacisl"]],
      .default = base_colors[["default"]]),
    # Store both rate-sorted and count-sorted versions
    data = map(data, function(df) {
      df %>%
        # arrange(desc(rate)) %>%
        mutate(name = label, y = rate, y2 = raw) %>%
        select(name, y, y2) %>%
        list_parse()}),
    data_count_sorted = map2(data, group, function(df, grp) {
      indicator_df %>%
        filter(group == grp, subgroup != "total") %>% 
        arrange(desc(raw)) %>%
        mutate(name = label, y = raw, y2 = rate) %>%
        select(name, y, y2) %>%
        list_parse()})) %>%
  select(id, type, color, data, data_count_sorted)

```


```{r}
##### Chart behavior: define JS to control behavior of drilldown (don't change) #####
### Drilldown js ----------------------
# define JS to control behavior or drilldown
drilldown_js <- paste0("function(e) {
  var chart = this;
  
  if (!e.seriesOptions) {
    return;
  }
  
  var seriesId = e.seriesOptions.id;
  
  if (!chart.originalDrilldownData) {
    chart.originalDrilldownData = {};
  }
  
  if (e.seriesOptions.data && e.seriesOptions.data_count_sorted) {
    chart.originalDrilldownData[seriesId] = {
      rateSorted: JSON.parse(JSON.stringify(e.seriesOptions.data)),
      countSorted: JSON.parse(JSON.stringify(e.seriesOptions.data_count_sorted))
    };
  }
  
  chart.yAxis[0].addPlotLine({
    id: 'total-line',
    value: ", total_rate, ",
    color: '", rc_black, "',
    width: 2,
    dashStyle: 'ShortDash',
    zIndex: 5,
    label: {style: {fontWeight: 'bold', fontSize: '0.8em'}, text: 'STATE AVERAGE: ", total_rate, "',
        align: 'right', rotation: 0, y: 5, x: -5, overflow: 'allow', crop: false}
  });
  
  chart.currentView = 'rates';
  
  chart.update({
    plotOptions: {
      bar: {
        states: { hover: { enabled: false } },
        dataLabels: {
          formatter: function() { return this.y; }
        }
      }
    }
  }, false);

  chart.yAxis[0].setTitle({ text: 'RATE (STATE AVERAGE: ", total_rate, ")' }, false);

  var backBtn = document.getElementById('hc-back-btn');
  if (backBtn) backBtn.style.display = 'block';
  
  var toggleBtn = document.getElementById('hc-toggle-btn');
  if (toggleBtn) {
    toggleBtn.style.display = 'block';
    toggleBtn.innerText = 'Show Counts';
  }
  
  chart.redraw();
  
  setTimeout(function() {
    requestAnimationFrame(function() {
      requestAnimationFrame(function() {
        resizeChart();
      });
    });
  }, 100);
}")


drillup_js <- paste0("function() { 
  var chart = this;
  chart.yAxis[0].removePlotLine('total-line');
  
  if (chart.originalDrilldownData && chart.options.drilldown && chart.options.drilldown.series) {
    chart.options.drilldown.series.forEach(function(drilldownSeries) {
      var seriesId = drilldownSeries.id;
      if (chart.originalDrilldownData[seriesId]) {
        drilldownSeries.data = chart.originalDrilldownData[seriesId].rateSorted;
      }
    });
  }
  
  var toggleBtn = document.getElementById('hc-toggle-btn');
  if (toggleBtn) toggleBtn.style.display = 'none';
  
  // ---- ADDED: hide custom back button ----
  var backBtn = document.getElementById('hc-back-btn');
  if (backBtn) backBtn.style.display = 'none';
  // ------------------------------------------
  
  chart.currentView = null;
  chart.yAxis[0].setTitle({ text: 'RATE' }, false);
  
  chart.update({
    plotOptions: {
      bar: {
        dataLabels: {
          formatter: function() { return this.y; }
        }
      }
    }
  }, false);
  
  chart.redraw();
  
  setTimeout(function() {
    requestAnimationFrame(function() {
      requestAnimationFrame(function() {
        resizeChart();
      });
    });
  }, 100);
}")

```


```{r}
##### Chart: Highcharts drilldown (don't change) #####
hc <- hchart(
  base_df,
  "bar",
  hcaes(x = group, y = rate, drilldown = group, y2 = raw, color=color),
  name = "Overview",
  colorByPoint = FALSE) %>%
  hc_chart(
    height = 400, 
    scrollablePlotArea = list(minHeight = 500, opacity = 0.9),
    events = list(
      drilldown = JS(drilldown_js),
      drillup = JS(drillup_js))) %>%
  hc_tooltip(enabled = FALSE) %>%
  hc_drilldown(
    allowPointDrilldown = TRUE,
    series = list_parse(drilldown_df),
    activeDataLabelStyle = list(
      color = rc_black,
      textDecoration = "none",
      fontWeight = "normal"
      ),
    activeAxisLabelStyle = list(
      color = rc_black,
      textDecoration = "none",
      cursor = "pointer"
      ),
    breadcrumbs = list(
    showFullPath = FALSE,
    buttonTheme = list(
      states = list(
        hover = list(fill = "none"),
        select = list(fill = "none")
      ),
      fill = "none",
      stroke = "none",
      `stroke-width` = 0,
      style = list(
        color = "transparent",
        fontSize = "0px"        # collapse text to nothing
      ),
      width = 0,
      height = 0,
      padding = 0
    ))
 ) %>%
  hc_title(
    text = chart_title,
    align = "center" ) %>% 
  hc_subtitle(
    text = case_when(
      race_grp == 'asian' ~ "Click the Asian or Total bar below to disaggregate data",
      race_grp == 'nhpi' ~ "Click the Pacific Islander or Total bar below to disaggregate data",
      .default = 'XXXXXXXXX'),
    align = "center") %>%
  hc_plotOptions(
    bar = list(
      dataLabels = list(
        enabled = TRUE,
        align = "right",
        inside = TRUE))) %>% 
  hc_xAxis(
    title=list(text=NULL),
    labels = list(
      enabled = TRUE,
      useHTML = TRUE,            # Required for complex CSS formatting (like word-break)
      formatter = JS("
        function() {
          // Explicitly wrap the text inside a fixed-width, wrapping div
          return '<div style=\"width: 100px; white-space: normal; word-wrap: break-word; text-align: right;\">' + this.value + '</div>';
        }
      "))) %>%
  hc_yAxis(
    title = list(text = "RATE"),
    labels=list(enabled=FALSE)) %>%# remove labels from rate axis) 
  hc_legend(enabled=FALSE) %>%
  hc_caption(
    text = paste("<b>Source:", data_source, "</b>", data_description, methodology_link),
    useHTML = TRUE,
    align = "left"
  ) %>%
  hc_add_theme(custom_theme) %>% 
  hc_responsive(
    rules = list(
      list(
        # Omit data_description on small screens
        condition = list(maxWidth = breakpoint_maxwidth),
        chartOptions = list(
          caption = 
            list(
              text = paste("<b>Source:</b>", data_source, methodology_link)
            )
        )
      )
    )
  )

htmltools::tagList(
  htmltools::tags$div(
    class = "chart-wrapper",
    style = "max-width: 1200px; width: 100%; margin: 0 auto; position: relative;",
  
  htmltools::tags$button(
    id = "hc-back-btn",
    htmltools::HTML("\u25C1 Back to Overview"),
    style = paste(
      "display: none;",
      "position: absolute;",
      "right: 10px;",
      "top: calc(50% - 60px);",   # above toggle button
      "width: 106px;",
      "white-space: normal;",
      "text-align: center;",
      paste0("background-color: ", rc_grey, ";"),
      "color: #333333;",
      "border: 1px solid #cccccc;",
      "border-radius: 3px;",
      "padding: 6px 8px;",
      "cursor: pointer;",
      "z-index: 10;",
      "line-height: 1.3;",
      sep = " "
    )
  ),
  
  htmltools::tags$button(
    id = "hc-toggle-btn",
    "Show Counts",
    style = paste(
      "display: none;",
      "position: absolute;",
      "right: 10px;",
      "top: calc(50% - 0px);",   # below back button
      "width: 106px;",
      "text-align: center;",
      paste0("background-color: ", rc_blue, ";"),
      paste0("color: ", rc_white, ";"),
      paste0("border: 1px solid ", rc_blue, ";"),
      "border-radius: 3px;",
      "padding: 6px 8px;",
      "cursor: pointer;",
      "z-index: 10;",
      "line-height: 1.3;",
      sep = " "
    )
  ),
  
  htmltools::tags$div(
    class = "chart-container",
    style = "height: auto; overflow: hidden; border: 1px solid #ccc;",
    hc
  )
),

htmltools::tags$script(HTML("
  document.getElementById('hc-back-btn').addEventListener('click', function() {
    if (typeof Highcharts !== 'undefined') {
      Highcharts.charts.forEach(function(chart) {
        if (chart) chart.drillUp();
      });
    }
  });

  document.getElementById('hc-toggle-btn').addEventListener('click', function() {
    var chart = Highcharts.charts.find(function(c) { return c; });
    if (!chart) return;
    
    var btn = document.getElementById('hc-toggle-btn');
    
    if (chart.currentView === 'rates') {
      chart.currentView = 'counts';
      btn.innerText = 'Show Rates';
      chart.yAxis[0].setTitle({ text: 'COUNT' }, false);
      chart.yAxis[0].removePlotLine('total-line');
      
      chart.series.forEach(function(s) {
        if (s.options.id && chart.originalDrilldownData[s.options.id]) {
          s.setData(chart.originalDrilldownData[s.options.id].countSorted, false);
        }
      });
      
      chart.update({
        plotOptions: {
          bar: {
            dataLabels: {
              formatter: function() { return this.y.toLocaleString('en-US'); }
            }
          }
        }
      }, false);
      
    } else {
      chart.currentView = 'rates';
      btn.innerText = 'Show Counts';
      chart.yAxis[0].setTitle({ text: 'RATE (STATE AVERAGE: ", total_rate, ")' }, false);
      
      chart.yAxis[0].addPlotLine({
        id: 'total-line',
        value: ", total_rate, ",
        color: '", rc_black, "',
        width: 2,
        dashStyle: 'ShortDash',
        zIndex: 5,
        label: {style: {fontWeight: 'bold', fontSize: '0.8em'}, text: 'STATE AVERAGE: ", total_rate, "',
        align: 'right', rotation: 0, y: 5, x: -5, overflow: 'allow', crop: false}
      });
      
      chart.series.forEach(function(s) {
        if (s.options.id && chart.originalDrilldownData[s.options.id]) {
          s.setData(chart.originalDrilldownData[s.options.id].rateSorted, false);
        }
      });
      
      chart.update({
        plotOptions: {
          bar: {
            dataLabels: {
              formatter: function() { return this.y; }
            }
          }
        }
      }, false);
    }
    
    chart.redraw();
  });

  
  function resizeChart() {
    if (typeof Highcharts !== 'undefined' && Highcharts.charts.length > 0) {
      Highcharts.charts.forEach(function(chart) {
        if (chart) {
          var wrapper = document.querySelector('.chart-wrapper');
          if (wrapper) {
            var targetWidth = wrapper.offsetWidth;
            chart.setSize(targetWidth, null, false);
          }
        }
      });
    }
  }
    
    setTimeout(function() {
      requestAnimationFrame(function() {
        requestAnimationFrame(resizeChart);
      });
    }, 100);
    window.addEventListener('resize', resizeChart);
  "))
)

```