---
title: "RACE COUNTS: Education"
author: "Authored by: Catalyst California"
date: "8/25/2025"
output:
  html_document:
    number_sections: no
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float: yes

subtitle: 'Draft Visuals'
---

<style type="text/css">
  h1{
  font-family: Rubik;
  }
  h2{
  font-family: Rubik;
  }
  h3{
  font-family: Rubik;
  }
  a{
  font-family: Rubik;
  }
  ul{
  font-family: Rubik;
  }
  ol{
  font-family: Rubik;
  }
  body{
  font-family: Rubik;
  font-size: 15px;
}
p{
  font-family: Rubik;
  font-size: 15px;
}
body { background-color: #FFFFFF; }

</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE)

#load the libraries via the below source as well as the credentials and set the theme
source("draft_visual_functions.R")

options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))

##Add RC theme##
rc_theme # load from the above draft_visual_functions.R script

##### UPDATE EACH YEAR AND FOR EACH ISSUE AREA ####
issue_name <- 'Education' # Issue names: Safety & Justice, Democracy, Economic Opportunity, Education, Healthy Built Environment, Health Care Access, Housing
issue <- 'educ' #Issue: crim, demo, educ, econ, hben, hlth, hous
rc_schema <- 'v7'
rc_yr <- '2025'
acs_yr <- '2023'

```

```{r echo=FALSE, include = FALSE}

####################### GET COUNTY DATA #####################################
county_list_query <- paste0("SELECT table_name FROM information_schema.tables WHERE table_type='BASE TABLE' AND table_schema='", rc_schema, "' AND table_name LIKE '%county_%", rc_yr, "' AND table_name LIKE '%", issue, "%' AND table_name NOT LIKE '%index%';")

county_list <- dbGetQuery(con, county_list_query) %>% pull(table_name) 

c_tables <- lapply(setNames(paste0("select * from ", rc_schema, ".", county_list), county_list), DBI::dbGetQuery, conn = con)

# create column with indicator name
c_tables <- map2(c_tables, names(c_tables), ~ mutate(.x, indicator = .y)) # create column with indicator name

# clean up and add columns
c_tables <- lapply(c_tables, function(x) x %>% 
  select(-any_of('geolevel'), -ends_with("_pop")) %>%
  mutate(issue = substring(indicator, 6,9),
         indicator = substring(indicator, 11),
         indicator = gsub(paste0('_county_', rc_yr), '', indicator),
         geolevel = 'county'))

# Get issue index
c_index_z <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_", issue, "_index_", rc_yr)) %>%
  rename(geo_name = county_name)  # make 'name' col generic

####################### POPULATION DATA #####################################
#Get total population
pop <- dbGetQuery(con, paste0("SELECT geoid, total_pop, geolevel FROM ", rc_schema, ".arei_race_multigeo WHERE geolevel = 'county'")) %>% select(-geolevel)

c_index_z <- left_join(c_index_z, pop, by = c("county_id" = "geoid"))

c_tables <- lapply(c_tables, function(x) 
  x %>% left_join(pop, by = c("county_id" = "geoid")))

## Make 'name' col generic
c_tables <- lapply(c_tables, function(x)
  x %>% rename(geo_name = county_name))

####################### GET STATE DATA #####################################
state_list_query <- paste0("SELECT table_name FROM information_schema.tables WHERE table_type='BASE TABLE' AND table_schema='", rc_schema, "' AND table_name LIKE '%state_%", rc_yr, "' AND table_name LIKE '%", issue, "%' AND table_name NOT LIKE '%index%';")

state_list <- dbGetQuery(con, state_list_query) %>% pull(table_name) 

s_tables <- lapply(setNames(paste0("select * from ", rc_schema, ".", state_list), state_list), DBI::dbGetQuery, conn = con)

# create column with indicator name
s_tables <- map2(s_tables, names(s_tables), ~ mutate(.x, indicator = .y)) # create column with indicator name

# clean up and add columns
s_tables <- lapply(s_tables, function(x) x %>% 
  select(-any_of('geolevel'), -ends_with("_pop")) %>%
  mutate(issue = substring(indicator, 6,9),
         indicator = substring(indicator, 11),
         indicator = gsub(paste0('_state_', rc_yr), '', indicator),
         geolevel = 'state'))


##Add 2025 v7 Source Data##
v7_sources <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_indicator_list_cntyst"))
v7_sources <- v7_sources %>% filter(v7_sources$arei_issue_area==issue_name) %>%
  select(arei_indicator, arei_issue_area, rate_descriptor, bar_chart_header, rate_rounding, methodology, data_source, api_name)


## Join metadata to data
c_tables <- lapply(c_tables, function(x) x %>% left_join(v7_sources, by = c("indicator" = "api_name")))
s_tables <- lapply(s_tables, function(x) x %>% left_join(v7_sources, by = c("indicator" = "api_name")))

names(c_tables)
names(s_tables)
```

<!-- *** -->
# Introduction
NOTE: The data and findings below are drafts subject to change and some pieces are still in the QA process.

RACE COUNTS provides a 3D view of racial equity:
<br>OUTCOME: How well people are doing. The higher the circle, the better the outcome.
<br>DISPARITY: How racial groups compare to one another. The further right the circle, the greater the differences by race.
<br>IMPACT: The total population. The bigger the circle, the larger the population.

<br><span style="color:purple">Purple</span> counties: Gains at Risk; 
<br><span style="color:orange">Orange</span> counties: Prosperity for the Few;
<br><span style="color:#c9cc14">Yellow</span> counties: Struggling to Prosper;
<br><span style="color:red">Red</span> counties: Stuck and Unequal.


# Race/Ethnicity Notes
<br>1) "Other" race includes those who identify with a race outside of the specifically named categories, such as Asian, White, etc.
<br>2) Race labels for bar charts: The "nh_" prefix signifies that a group is non-Latinx (excludes Latinx).
<br>3) The "twoormor" group represents those who identify as Two or More Races.

# Key Findings

* There are several key areas where racial disparities exist in educational outcomes across different racial and ethnic groups in the state of California.

* Across the state, NH-AIAN, NH-Black, Latinx, and NH-Pacific Islander are more likely to experience chronic absenteeism than other groups. (See State Barchart on Chronically Absent Students (%))

* Additionally, NH-AIAN, NH-Black, Latinx, and NH-Pacific Islander students have lower graduation rates than the state average. NH-Black student graduation rates are 16.7 percentage points lower than the group with the highest rate. (See State Barchart on Graduates as a Percentage of Four-Year Adjusted Cohort Students)

* NH- Filipinx, NH- Asian, NH- White, and NH- Students of Two or More races experience lower suspension rates. (See State Barchart on students Suspended per 100 Students)

* More than 50% of Asian, Filipinx, Two or More Races, and White third graders scored proficient in English Language Arts, while less than a third of NH-PI, NH-Black, NH-AIAN, and Latinx students score proficient or better. (See State Barchart on 3rd Graders Scoring Proficient or Better in English Language Arts (%) and 3rd Graders Scoring Proficient or Better in Math (%))

* ECE access rates are low across the state for all groups, even the best rate means less than one in two children has acess to licensed ECE programs. Still NH-AIAN, Latinx, and NH-Black children all have less acccess to licensed ECE programs than the average child in California. (See State Barchart on Children 0-5 Enrolled in Licensed ECE Program (%))

# Indicators 

## Education Index - UPDATED
```{r, echo=FALSE}
# Remove counties without quadrant values. Cap index perf/disp z-scores at 2 and -2. More info: https://advancementproject.sharepoint.com/:w:/s/Portal/EX59kBOn8iRNrLuY1Sfk3JABT34dO3sj1j9fwkuUxLqUgQ?e=bGyEaZ
  c_index_z <- c_index_z %>% mutate(arei_issue_area=issue_name)
  c_index_chart <- index_scatterplot(c_index_z, 2)

  c_index_chart
```

* San Francisco has among the highest racial disparities, driven by large inequities in chronic absenteeism and graduation rates.

* Most Northern California region counties have lower than average outcome, with many counties also exhibiting higher levels of disparity. 




## Chronic Absenteeism - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}

c_1_chart <- county_scatterplot(c_tables[[1]])
c_1_chart
```
* San Francisco County is by far the most racially disparate on chronic absenteeism. NHPI students have the highest absenteeism rate at 65.9%, which is  58.2 percentage points higher than the group with the lowest rate.

* Plumas County has the highest rates of absenteeism, with more than half of students (56.2%) being chronically absent. All groups have higher rates, however, Latinx students are the most likely to be chronically absent (61.8%).

### State Barchart
```{r, echo=FALSE}
  s_1_ <- state_barchart(s_tables[[1]])
  s_1_
``` 
* Statewide, NHPI, NH-Black, NH-AIAN, and Latinx students have the highest rates for chronic absenteeism ranging from 28.9% to 36.6% respectively. 

* NH-Black, NHPI, and NH-AIAN students are 3.5 times more likely to be chronically absent than the group with the lowest rate.

### County Barchart

```{r, echo=FALSE}
c_1_long <- county_barchart(c_tables[[1]]) 
c_1_long

```


## High School Graduation - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_2_chart <- county_scatterplot(c_tables[[2]])
c_2_chart
```
* Mono, Nevada, Inyo, and San Francisco Counties have the lowest graduation rates and the highest racial disparities on this measure. 

* In Mono, only one in three Latinx students graduate.

### State Barchart
```{r, echo=FALSE}
  s_2_ <- state_barchart(s_tables[[2]])
  s_2_

```
* NH-AIAN, NH-Black, Latinx, and NH-PI students have below average graduation rates. Black and NH-AIAN student graduation rates are roughly 16 percentage points lower than the group with the highest rate.

### County Barchart

```{r, echo=FALSE}
c_2_long <- county_barchart(c_tables[[2]]) 
c_2_long

```


## Third Grade English Proficiency - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_3_chart <- county_scatterplot(c_tables[[3]])
c_3_chart
```
* Modoc County has the lowest rates in this measure, with only one in five 3rd graders scoring Proficient or Better in ELA.

* Imperial County is the most disparate in this measure, with a 41.5 percentage point difference between the groups with the highest and lowest rate. 

 

### State Barchart

```{r, echo=FALSE}
  s_3_ <- state_barchart(s_tables[[3]])
  s_3_
``` 
* Statewide, two groups emerge. In the first group, fewer than one in three NH-AIAN, Latinx, NH-Black, and NHPI students scored proficient. While in the second group, more than 50% of Asian, Filipinx, Two or More Races, and White third graders scored proficient in English Language Arts. 


### County Barchart

```{r, echo=FALSE}
c_3_long <- county_barchart(c_tables[[3]]) 
c_3_long
```


## Third Grade Math Proficiency - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_4_chart <- county_scatterplot(c_tables[[4]])
c_4_chart
```
* Modoc County is the lowest performing on this measure with less than 25% of its students scoring Proficient or Better in Math.

* Larger population counties have above average levels of racial disparities among racial and ethnic groups for this indicator.

### State Barchart

```{r, echo=FALSE}
  s_4_ <- state_barchart(s_tables[[4]])
  s_4_
```
* Statewide, two groups emerge. In the first group, about one in three NH-AIAN, NH-Black, Latinx, and NH-PI students scored proficient. While in the second group, more than 50% of Asian, Filipinx, White, Two or More Races third graders scored proficient in math. 

### County Barchart

```{r, echo=FALSE}
c_4_long <- county_barchart(c_tables[[4]]) 
c_4_long
```


## Suspensions - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_5_chart <- county_scatterplot(c_tables[[5]])
c_5_chart
```
* Inyo County is the most racially disparate for this measure, with suspension rates for NH-AIAN students reaching 7.1% while the suspension rates for the lowest group is 0.

* Three counties (Modoc, Lake, and Del Norte) in the Northern California region are the lowest performing for this measure, with suspension rates at or nearing 10%. 


### State Barchart

```{r, echo=FALSE}
  s_5_ <- state_barchart(s_tables[[5]])
  s_5_
```
* NH-Pacific Islander, Latinx, NH-AIAN, and NH-Black students are suspended at above state average rates. 

* NH-Black students are suspended at a rate 8.0 times higher than the lowest suspension rate. 

### County Barchart

```{r, echo=FALSE}
c_5_long <- county_barchart(c_tables[[5]]) 
c_5_long
```


## ECE Access - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_6_chart <- county_scatterplot(c_tables[[6]])
c_6_chart
```
* 22.2% of San Bernardino County children ages 0-5 have access to a licensed ECE program, which is a 62.4 percentage point difference from Modoc County (84.6%), the county with the most access. 



### State Barchart

```{r, echo=FALSE}
  s_6_ <- state_barchart(s_tables[[6]])
  s_6_
``` 
* ECE access rates are low across the state and for all groups, even the best rate means less than one in two children has access. Still NH-AIAN, Latinx, and NH-Black children all have less ECE access than the state average.

### County Barchart

```{r, echo=FALSE}
c_6_long <- county_barchart(c_tables[[6]]) 
c_6_long
```



## Diversity of Teachers  - NOT UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_7_chart <- county_scatterplot(c_tables[[7]])
c_7_chart
```
* Larger population counties overall are in the Yellow (Lower disparity, Lower outcome) Quadrant, with Low Angeles County having the lowest disparity.

* All but two (Mariposa and Tuolumne) Central Valley counties are in the Yellow (Lower disparity, Lower outcome) and Red (Higher disparity, Lower outcome) Quadrants.

### State Barchart

```{r, echo=FALSE}
  s_7_ <- state_barchart(s_tables[[7]])
  s_7_
``` 
* White students have by far the highest representation, with more than 13 White teachers and staff per 100 White students. Compare this to NH-AIANs, the group with the next highest rate, at five per 100 students.

### County Barchart

```{r, echo=FALSE}
c_7_long <- county_barchart(c_tables[[7]]) 
c_7_long
```

```{r, include=FALSE}
##Disconnect from db
dbDisconnect(con)
dbDisconnect(con2)

```


```{r echo=FALSE, include = FALSE}
# below is a javascript chunk for the county bar-charts. The chunk automatically selects the 'Alameda' county as the first chunk. Without the chunk below, as you can see from the output above, the first bar-chart shown sums the race/ethnicity rates for all of the counties. 
# Each line that starts with document is for each graph (0-5) for a total of 6 graphs. If another graph is added, then you would simply add another line and label it as [6]. 

# Notice that it isn't an R chunk, it is a JS chunk. :) 
```


```{js}
function filter_default() {
    document.getElementsByClassName("selectized") 
[0].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[1].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[2].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[3].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[4].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[5].selectize.setValue("Alameda", false);

 }
window.onload = filter_default;
```




