---
title: "RACE COUNTS: Health Care Access               "
author: "Authored by: Catalyst California"
date: "8/25/2025"
output:
  html_document:
    number_sections: no
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float: yes

subtitle: 'Draft Visuals'
---
  
<style type="text/css">
  h1{
  font-family: Rubik;
  }
  h2{
  font-family: Rubik;
  }
  h3{
  font-family: Rubik;
  }
  a{
  font-family: Rubik;
  }
  ul{
  font-family: Rubik;
  }
  ol{
  font-family: Rubik;
  }
  body{
  font-family: Rubik;
  font-size: 15px;
}
p{
  font-family: Rubik;
  font-size: 15px;
}
body { background-color: #FFFFFF; }

</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE)

#load the libraries via the below source as well as the credentials and set the theme
source("W:/Project/RACE COUNTS/2025_v7/RC_Github/AB/RaceCounts/Functions/draft_visual_functions.R")

options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))

##Add RC theme##
rc_theme # load from the above draft_visual_functions.R script
issue <- 'Health Care Access' # 7 options: Crime & Justice, Democracy, Economic Opportunity, Education, Healthy Built Environment, Health Care Access, Housing
rc_schema <- 'v7'
rc_yr <- '2025'
acs_yr <- '2023'

```

```{r echo=FALSE, include = FALSE}
####################### ADD COUNTY DATA #####################################

c_1 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_got_help_county_", rc_yr))
c_2 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_health_insurance_county_", rc_yr))
c_3 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_life_expectancy_county_", rc_yr))
c_4 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_low_birthweight_county_", rc_yr))
c_5 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_usual_source_of_care_county_", rc_yr))
c_6 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_preventable_hospitalizations_county_", rc_yr)) %>% 
  select(-total_pop) #select is just needed for this indicator


c_index_z <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_index_", rc_yr))

####################### POPULATION DATA #####################################
#Join total population to index view, food access, and asthma data frames
pop <- dbGetQuery(con2, paste0("SELECT geoid, dp05_0001e FROM demographics.acs_5yr_dp05_multigeo_", acs_yr)) %>% rename(total_pop = dp05_0001e)

#c_index_pctile <- left_join(x = c_index_pctile, y = pop, by = c("county_id" = "geoid"))
c_index_z <- left_join(x = c_index_z, y = pop, by = c("county_id" = "geoid"))

c_1 <- left_join(x = c_1, y = pop, by = c("county_id" = "geoid"))
c_3 <- left_join(x = c_3, y = pop, by = c("county_id" = "geoid"))
c_4 <- left_join(x = c_4, y = pop, by = c("county_id" = "geoid"))
c_5 <- left_join(x = c_5, y = pop, by = c("county_id" = "geoid"))
c_6 <- left_join(x = c_6, y = pop, by = c("county_id" = "geoid"))


####################### ADD STATE DATA #####################################
s_1 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_got_help_state_", rc_yr))
s_2 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_health_insurance_state_", rc_yr))
s_3 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_life_expectancy_state_", rc_yr))
s_4 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_low_birthweight_state_", rc_yr))
s_5 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_usual_source_of_care_state_", rc_yr))
s_6 <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_hlth_preventable_hospitalizations_state_", rc_yr))

##Add 2025 v7 Source Data##
v7_sources <- dbGetQuery(con, paste0("SELECT * FROM ", rc_schema, ".arei_indicator_list_cntyst"))
v7_sources <- v7_sources %>% filter(v7_sources$arei_issue_area==issue)

# merge with urban type
# region_urban_type <- dbGetQuery(con, paste0("SELECT geo_id AS county_id, region, urban_type FROM ", rc_schema, ".arei_multigeo_list")
# c_1 <- left_join(c_1, region_urban_type)
# c_2 <- left_join(c_2, region_urban_type)
# c_3 <- left_join(c_3, region_urban_type)
# c_4 <- left_join(c_4, region_urban_type)
# c_5 <- left_join(c_5, region_urban_type)
# c_6 <- left_join(c_6, region_urban_type)
```

<!-- *** -->
# Introduction
NOTE: The data and findings below are drafts subject to change and some pieces are still in the QA process.

RACE COUNTS provides a 3D view of racial equity:
<br>PERFORMANCE: How well people are doing. The higher the circle, the better the performance.
<br>DISPARITY: How racial groups compare to one another. The further right the circle, the greater the differences by race.
<br>IMPACT: The total population. The bigger the circle, the larger the population.

<br><span style="color:purple">purple</span> counties: Gains at Risk; 
<br><span style="color:orange">Orange</span> counties: Prosperity for the Few;
<br><span style="color:#c9cc14">Yellow</span> counties: Struggling to Prosper;
<br><span style="color:red">Red</span> counties: Stuck and Unequal.


# Race/Ethnicity Notes
<br>1) "Other" race includes those who identify with a race outside of the specifically named categories, such as Asian, White, etc.
<br>2) Race labels for bar charts: The "nh_" prefix signifies that a group is non-Latinx (excludes Latinx).
<br>3) The "twoormor" group represents those who identify as Multiracial.

# Indicators

## Health Care Access Index - UPDATED
```{r, echo=FALSE}
# Remove counties without quadrant values. Cap index perf/disp z-scores at 2 and -2. More info: https://advancementproject.sharepoint.com/:w:/s/Portal/EX59kBOn8iRNrLuY1Sfk3JABT34dO3sj1j9fwkuUxLqUgQ?e=bGyEaZ
  c_index_z <- c_index_z %>% mutate(arei_issue_area=v7_sources$arei_issue_area[1])
  c_index_chart <- index_scatterplot(c_index_z, 2)

  c_index_chart
```

* 
* 
* 
* 

## Got Help - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE} 

c_1 <- c_1 %>% mutate(arei_indicator=v7_sources$arei_indicator[2], bar_chart_header=v7_sources$bar_chart_header[2], description = v7_sources$methodology[2])
c_1_chart <- county_scatterplot(c_1)
c_1_chart

```
* 
* 
* 

### State Barchart
```{r, echo=FALSE}

  title = v7_sources$bar_chart_header[2]
  s_1_ <- state_barchart(s_1, title)
  s_1_

``` 
* 
* 

### County Barchart

```{r, echo=FALSE}
c_1_long <- county_barchart(c_1, title) 
c_1_long
```



## Lack of Health Insurance - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_2 <- c_2 %>% mutate(arei_indicator=v7_sources$arei_indicator[1], bar_chart_header=v7_sources$bar_chart_header[1], description = v7_sources$methodology[1])
c_2_chart <- county_scatterplot(c_2)
c_2_chart

```
* Similar to the mental health/substance abuse help indicator, Marin County has the third highest overall rate of health insurance, but the second highest racial disparity. Latinx residents of the county are 7x more likely to be uninsured rate than White residents.
* Sierra County has the highest disparity and third-worst performance of any county in the state. Over half of its Latinx population is uninsured -- a rate more than 10x higher than its white population. 
* Los Angeles County is one of the worst performing counties in the state in terms of health insurance rates among its population. 9% of all its residents are uninsured, including 13.2% of its Latinx residents.


### State Barchart
```{r, echo=FALSE}
  title = v7_sources$bar_chart_header[1]
  s_2_ <- state_barchart(s_2, title)
  s_2_

```
* White residents are about 2 times less likely to be uninsured than the average Californian.
* Latinx and American Indian/Alaska Native Californians are about 3 times more likely to be uninsured than White residents. 
* Californians who identify with a racial group under the category of "other" are the most likely to be uninsured. Although it is difficult to know what types of individuals are identifying in this group, it likely includes many who are Latinx (as "Another Race" includes Latinx for this indicator) and those with origins in the Middle East and North Africa, among others.


### County Barchart

```{r, echo=FALSE}
c_2_long <- county_barchart(c_2, title) 
c_2_long
```


## Life Expectancy - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_3 <- c_3 %>% mutate(arei_indicator=v7_sources$arei_indicator[5], bar_chart_header=v7_sources$bar_chart_header[5], description = v7_sources$methodology[5])
c_3_chart <- county_scatterplot(c_3)
c_3_chart
```
* Five of the six counties in Southern California have greater than average racial disparities in life expectancy. 
* Every county in the San Joaquin Valley has below average life expectancy rates.
* Less populous, less urban counties tend to have fewer racial disparities in life expectancy. Of the 15 least disparate counties in this measure, none of them are classified as urban and only one is among the top ten most populous counties in the state.
* The state's most populous, more urban counties tend to exhibit greater racial disparities in life expectancy. San Francisco, San Bernardino, Santa Clara, Los Angeles, San Diego, and Alameda counties are all among the top 20 for disparity levels in this metric.

### State Barchart

```{r, echo=FALSE}

  title = v7_sources$bar_chart_header[5]
  s_3_ <- state_barchart(s_3, title)
  s_3_

``` 
* Asians have the highest life expectancy, followed by non-Latinx White Californians. Non-Latinx Blacks have the lowest life expectancy rate in the state, six years lower than the average statewide rate.

### County Barchart

```{r, echo=FALSE}
c_3_long <- county_barchart(c_3, title) 
c_3_long
```


## Low Birthweight - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_4 <- c_4 %>% mutate(arei_indicator=v7_sources$arei_indicator[4], bar_chart_header=v7_sources$bar_chart_header[4], description = v7_sources$methodology[4])
c_4_chart <- county_scatterplot(c_4)
c_4_chart
```
* There is no data for 17 of the 20 counties in the Northern/Sierra region for this measure.
* There is also a moderately strong, negative correlation between population and performance. Higher population counties tend to experience higher rates of infants with low birthweight. The notable exception is Orange County, which is the third most populous county in the state and is firmly in the purple quadrant (higher performance, lower disparity).
* San Francisco County and Kings County exhibit by far the largest racial disparities in rates of low birthweight, with Black birth outcomes notably on the wrong end of these disparities. In SF and Kings respectively, Black infants are about 3x and 2x more likely to be underweight than white infants.
* The three counties with the lowest disparity (Imperial, Humboldt, and Napa) do not have data for Black births. Given the stark disparities in Black birth outcomes, this is at least one reason why disparity is so low in those counties.


### State Barchart

```{r, echo=FALSE}

  title = v7_sources$bar_chart_header[4]
  s_4_ <- state_barchart(s_4, title)
  s_4_

```
* Black babies have the significantly highest likelihood of having low birthweight of any racial group in California. They are 50% more likely than the group with the next highest rate (American Indians and Alaska Natives), and more than 2x more likely than the group with the lowest rate (white Californians).


### County Barchart

```{r, echo=FALSE}
c_4_long <- county_barchart(c_4, title) 
c_4_long
```


## Usual Source of Care - UPDATED {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_5 <- c_5 %>% mutate(arei_indicator=v7_sources$arei_indicator[3], bar_chart_header=v7_sources$bar_chart_header[3], description = v7_sources$methodology[3])
c_5_chart <- county_scatterplot(c_5)
c_5_chart

```
* All nine Bay Area counties have above average usual source of care rates among their populations and six counties are in the purple Quadrant (Lower Disparity, Higher Performance).
* Counties with larger populations tend to have lower levels of racial disparity in their usual source of care rates, with none of the 16 counties with 500,000 or more residents appearing in the higher disparity quadrants (Orange and Red).

### State Barchart

```{r, echo=FALSE}

  title = v7_sources$bar_chart_header[3]
  s_5_ <- state_barchart(s_5, title)
  s_5_

```
* Latinx, Pacific Islander, AIAN, and Asian Californians are less likely to have a usual source of care than the average Californian.
* Latinx Californians are over nine percentage points less likely to have a usual source of care than White Californians.


### County Barchart

```{r, echo=FALSE}

c_5_long <- county_barchart(c_5, title) 
c_5_long

```


## Preventable Hospitalizations {.tabset .tabset-fade}

### Scatterplot

```{r, echo=FALSE}
c_6 <- c_6 %>% mutate(arei_indicator=v7_sources$arei_indicator[6], bar_chart_header=v7_sources$bar_chart_header[6], description = v7_sources$methodology[6])
c_6_chart <- county_scatterplot(c_6)
c_6_chart

```
* The four highest levels of racial disparity in the state, in Imperial, Amador, Lassen, and Mariposa counties, all represent extremely values compared to the rest of the state. In the 2nd-4th highest counties, their small population sizes (>40,000) could help inflate these measures.
* In Imperial County, by far the most racially disparate county in this measure, white residents have by far the worst outcomes in preventable hospitalizations. They are hospitalized at around 5x the rate of Asian residents and nearly twice the rate of Black and Latinx residents in the county.
* In many counties, some of the highest rates are for Another Race, local knowledge will be needed to better understand who makes up this group in each county.

### State Barchart

```{r, echo=FALSE}

  title = v7_sources$bar_chart_header[6]
  s_6_ <- state_barchart(s_6, title)
  s_6_

``` 
* The Black preventable hospitalization rate is nearly 5 times higher than that of the group with the lowest rate, and 2.5 times higher than even the state average.
* The preventable hospitalization rate for white Californians is about 10% higher than the overall state average and more than 2x higher than the racial group with the lowest rate. This is the only health metric in which white residents experience worse outcomes than the state average.


### County Barchart

```{r, echo=FALSE}

c_6_long <- county_barchart(c_6, title) 
c_6_long

```


```{r echo=FALSE, include = FALSE}
# below is a javascript chunk for the county bar-charts. The chunk automatically selects the 'Alameda' county as the first chunk. Without the chunk below, as you can see from the output above, the first bar-chart shown sums the race/ethnicity rates for all of the counties. 
# Each line that starts with document is for each graph (0-5) for a total of 6 graphs. If another graph is added, then you would simply add another line and label it as [6]. 

# Notice that it isn't an R chunk, it is a JS chunk. :) 
```

```{r, echo=FALSE}
##Closing database connections##
dbDisconnect(con)
dbDisconnect(con2)

```

```{js}
function filter_default() {
    document.getElementsByClassName("selectized") 
[0].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[1].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[2].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[3].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[4].selectize.setValue("Alameda", false);

 document.getElementsByClassName("selectized") 
[5].selectize.setValue("Alameda", false);

 }
window.onload = filter_default;
```





