---
title: "RC v7 City Index"
author: "Authored by: Catalyst California"
date: "8/25/2025"
output:
  html_document:
    number_sections: no
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float: yes
subtitle: 'Draft Visuals'
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<style type="text/css">
  h1{
  font-family: Rubik;
  }
  h2{
  font-family: Rubik;
  }
  h3{
  font-family: Rubik;
  }
  a{
  font-family: Rubik;
  }
  ul{
  font-family: Rubik;
  }
  ol{
  font-family: Rubik;
  }
  body{
  font-family: Rubik;
  font-size: 15px;
}
p{
  font-family: Rubik;
  font-size: 15px;
}

</style>



### RC 2025: City Index Table

```{r, echo=FALSE, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE)

# set up workspace
packages <- c("DT", "RPostgres", "formattable", "knitr", "dplyr", "htmltools", "tidyr", "stringr", "scales", "colorspace", "sf")  

install_packages <- packages[!(packages %in% installed.packages()[,"Package"])] 

if(length(install_packages) > 0) { 
  install.packages(install_packages) 
  
} else { 
  
  print("All required packages are already installed.") 
} 

for(pkg in packages){ 
  library(pkg, character.only = TRUE) 
} 

# load PostgreSQL driver
source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("racecounts")

# setwd("W:/Project/RACE COUNTS/2025_v7/Visualizations/") 

#### Data Prep Steps ####

### Load index data ###
city_index_ <- dbGetQuery(con, paste0("SELECT * FROM v7.arei_multigeo_list WHERE geolevel = 'place'")) 

# drop unneeded cols and cities w/o index scores
city_index <- city_index_ %>% select(-c(geoid, county_id, urban_type, geolevel, nh_other_pop, nh_twoormor_pop, starts_with("pct_"), ends_with("_disparity_z"), ends_with("_performance_z"), ends_with("_disparity_rank"), ends_with("_performance_rank")))
city_index <- city_index %>% filter(!is.na(disparity_z))

# round z-scores
   city_index <- mutate(city_index,
   disparity_z = round(disparity_z,2),
   performance_z = round(performance_z,2),
   )
    
# Calc quadrants
city_index <- city_index
  city_index$quadrant =
    ifelse(city_index$performance_z < 0 & city_index$disparity_z > 0, 'red',
           ifelse(city_index$performance_z >= 0 & city_index$disparity_z >= 0, 'orange',
                  ifelse(city_index$performance_z >= 0 & city_index$disparity_z < 0, 'purple',
                         ifelse(city_index$performance_z < 0 & city_index$disparity_z <= 0, 'yellow', NA))))
   
# order table by disparity_rank and reorder cols
city_index <- arrange(city_index, -desc(disparity_rank))
city_index <- city_index %>% select(geo_name, region, disparity_rank, disparity_z, performance_rank, performance_z, quadrant, everything())

## table ##
datatable(city_index,
class = 'cell-border stripe',
colnames=c(
'City'= 'geo_name',
'Disparity Rank'= 'disparity_rank',
'Disparity Z-Score'= 'disparity_z',
'Outcome Rank'= 'performance_rank',
'Outcome Z-Score'= 'performance_z',
'Category' = 'quadrant',
'Population'= 'total_pop',
'Region'= 'region',
'AIAN Population'= 'aian_pop',
'Asian Population' = 'nh_asian_pop',
'Black Population' = 'nh_black_pop',
'Latinx Population'= 'latino_pop',
'NHPI Population'= 'pacisl_pop',
'White Population' = 'nh_white_pop'
 ),
style ='default',
rownames = FALSE,
filter ='top',
extensions = c(
  # 'Scroller',
  # 'FixedHeader',
               'Buttons',
               'KeyTable', 'FixedColumns'),

options = list(
  dom = 'Bfrtip',
  buttons = c('csv','excel'),
pageLength = 15,
autoWidth = TRUE,
scrollX = T,
scroller = TRUE
  # columnDefs = list(list(width = '200px', targets = c(1,23))),
#, searchHighlight = TRUE
)
,
caption = htmltools::tags$caption(
style = 'caption-side: bottom; text-align: left;',
htmltools::strong('Z-scores range from -1 to 1. A higher disparity rank means a higher level of disparity. A higher outcome rank means better overall outcomes.'))
)


```

```{r, echo=FALSE, cache=TRUE}
## NOTE: Data viz below based on: W:\Project\RACE COUNTS\2022_v4\Visualizations\RC_2022_Indexes.RMD ###

knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE)

#Set up workspace
packages <- c("DT", "RPostgres", "formattable","knitr","dplyr","htmltools","tidyr","stringr","scales","colorspace","sf", "leaflet",
"rgdal","rpostgis","RColorBrewer","ggplot2","scales","highcharter","here","htmltools","devtools","broom","kableExtra","reshape2")  

install_packages <- packages[!(packages %in% installed.packages()[,"Package"])] 

if(length(install_packages) > 0) { 
  install.packages(install_packages) 
  
} else { 
  
  print("All required packages are already installed.") 
} 

for(pkg in packages){ 
  library(pkg, character.only = TRUE) 
} 

options(scipen=999)
options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))


####################### ADD RC THEME #####################################
rc_theme <- hc_theme(
    # race counts colors
  colors = c("orange", "#800080", "red", "yellow"),
  chart = list(
    backgroundColor = "#F7F7F7", # RC white
    style = list(
          fontFamily = "Rubik") 
  ),
  title = list(
    style = list(
      color = "#070024",
      fontFamily = "Rubik")
  ),
  subtitle = list(
    style = list(
      color = "#8E8C8F", # medium grey
      fontFamily = "Rubik")
  ),
  caption = list(
    style = list(
      color = "#8E8C8F",
      fontFamily = "Rubik")
  ),
    axis = list(
    style = list(
      color = "#8E8C8F",
      fontFamily = "Rubik")
  ),
  legend = list(
    itemStyle = list(
      fontFamily = "Rubik",
      color = "#070024"),
    itemHoverStyle = list(
       fontFamily = "Rubik",
      color = "#070024")
  ))

# Add commas to long numbers
lang <- getOption("highcharter.lang")
lang$thousandsSep <- ","
options(highcharter.lang = lang)

```


## Composite Index - UPDATED

```{r}
# create scatterplot
c_chart_index <- city_index %>% 
    hchart("scatter", 
    hcaes(x=disparity_z, y=performance_z, size=total_pop, group=quadrant),
      tooltip = 
         list(pointFormat = 
                "<strong>City: </strong>{point.geo_name} <br><strong>Outcome Rank: </strong>{point.performance_rank} <br><strong>Disparity Rank: </strong>{point.disparity_rank} <br><strong>Population: </strong>{point.total_pop:,.0f}")) %>%
  hc_title(text = "Composite Index") %>%
  hc_xAxis(title = list(text = "Disparity Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_yAxis(title = list(text = "Outcome Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_caption(
    text = "Data source: Various",
    align = "center"
     ) %>% hc_size(height=700,width=700)

  hc_add_theme(c_chart_index, rc_theme) 

```




## Composite Index - Bay Area

```{r}
# Filter for specified region 
city_index_b <- city_index %>% filter(region == 'Bay Area Counties')

c_chart_index <- city_index_b %>% 
    hchart("scatter", 
    hcaes(x=disparity_z, y=performance_z, size=total_pop, group=quadrant),
      tooltip = 
         list(pointFormat = 
                "<strong>City: </strong>{point.geo_name} <br><strong>Outcome Rank: </strong>{point.performance_rank} <br><strong>Disparity Rank: </strong>{point.disparity_rank} <br><strong>Population: </strong>{point.total_pop:,.0f}")) %>%
  hc_title(text = "Composite Index: Bay Area") %>%
  hc_xAxis(title = list(text = "Disparity Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_yAxis(title = list(text = "Outcome Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_caption(
    text = "Data source: Various",
    align = "center"
     ) %>% hc_size(height=700,width=700)

  hc_add_theme(c_chart_index, rc_theme) 

```



## Composite Index - Southern California

```{r}
# Filter for specified region
city_index_s <- city_index %>% filter(region == 'Southern California')

c_chart_index <- city_index_s %>% 
    hchart("scatter", 
    hcaes(x=disparity_z, y=performance_z, size=total_pop, group=quadrant),
      tooltip = 
         list(pointFormat = 
                "<strong>City: </strong>{point.geo_name} <br><strong>Outcome Rank: </strong>{point.performance_rank} <br><strong>Disparity Rank: </strong>{point.disparity_rank} <br><strong>Population: </strong>{point.total_pop:,.0f}")) %>%
  hc_title(text = "Composite Index: Southern California") %>%
  hc_xAxis(title = list(text = "Disparity Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_yAxis(title = list(text = "Outcome Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_caption(
    text = "Data source: Various",
    align = "center"
     ) %>% hc_size(height=700,width=700)

  hc_add_theme(c_chart_index, rc_theme) 

```


## Composite Index - San Joaquin Valley

```{r}
# Filter for specified region
city_index_sjv <- city_index %>% filter(region == 'San Joaquin Valley')

c_chart_index <- city_index_sjv %>% 
    hchart("scatter", 
    hcaes(x=disparity_z, y=performance_z, size=total_pop, group=quadrant),
      tooltip = 
         list(pointFormat = 
                "<strong>City: </strong>{point.geo_name} <br><strong>Outcome Rank: </strong>{point.performance_rank} <br><strong>Disparity Rank: </strong>{point.disparity_rank} <br><strong>Population: </strong>{point.total_pop:,.0f}")) %>%
  hc_title(text = "Composite Index: San Joaquin Valley") %>%
  hc_xAxis(title = list(text = "Disparity Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_yAxis(title = list(text = "Outcome Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_caption(
    text = "Data source: Various",
    align = "center"
     ) %>% hc_size(height=700,width=700)

  hc_add_theme(c_chart_index, rc_theme) 

```


## Composite Index - Central Coast

```{r}
# Filter for specified region
city_index_cc <- city_index %>% filter(region == 'Central Coast Counties')

c_chart_index <- city_index_cc %>% 
    hchart("scatter", 
    hcaes(x=disparity_z, y=performance_z, size=total_pop, group=quadrant),
      tooltip = 
         list(pointFormat = 
                "<strong>City: </strong>{point.geo_name} <br><strong>Outcome Rank: </strong>{point.performance_rank} <br><strong>Disparity Rank: </strong>{point.disparity_rank} <br><strong>Population: </strong>{point.total_pop:,.0f}")) %>%
  hc_title(text = "Composite Index: Central Coast") %>%
  hc_xAxis(title = list(text = "Disparity Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_yAxis(title = list(text = "Outcome Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_caption(
    text = "Data source: Various",
    align = "center"
     ) %>% hc_size(height=700,width=700)

  hc_add_theme(c_chart_index, rc_theme) 

```


## Composite Index - Sacramento Area

```{r}
# Filter for specified region
city_index_sac <- city_index %>% filter(region == 'Sacramento Area')

c_chart_index <- city_index_sac %>% 
    hchart("scatter", 
    hcaes(x=disparity_z, y=performance_z, size=total_pop, group=quadrant),
      tooltip = 
         list(pointFormat = 
                "<strong>City: </strong>{point.geo_name} <br><strong>Outcome Rank: </strong>{point.performance_rank} <br><strong>Disparity Rank: </strong>{point.disparity_rank} <br><strong>Population: </strong>{point.total_pop:,.0f}")) %>%
  hc_title(text = "Composite Index: Sacramento Area") %>%
  hc_xAxis(title = list(text = "Disparity Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_yAxis(title = list(text = "Outcome Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_caption(
    text = "Data source: Various",
    align = "center"
     ) %>% hc_size(height=700,width=700)

  hc_add_theme(c_chart_index, rc_theme) 

```


## Composite Index - Northern / Sierra Counties

```{r}
# Filter for specified region
city_index_ns <- city_index %>% filter(region == 'Northern / Sierra Counties')

c_chart_index <- city_index_ns %>% 
    hchart("scatter", 
    hcaes(x=disparity_z, y=performance_z, size=total_pop, group=quadrant),
      tooltip = 
         list(pointFormat = 
                "<strong>City: </strong>{point.geo_name} <br><strong>Outcome Rank: </strong>{point.performance_rank} <br><strong>Disparity Rank: </strong>{point.disparity_rank} <br><strong>Population: </strong>{point.total_pop:,.0f}")) %>%
  hc_title(text = "Composite Index: Northern / Sierra Counties") %>%
  hc_xAxis(title = list(text = "Disparity Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_yAxis(title = list(text = "Outcome Z-Score"), max = 1.2, min = -1.2, tickInterval = .5) %>% 
  hc_caption(
    text = "Data source: Various",
    align = "center"
     ) %>% hc_size(height=700,width=700)

  hc_add_theme(c_chart_index, rc_theme) 

```


<!-- ```{r, cache=TRUE}

##### City Index Map ####
# set up workspace
library(DT)
library(RPostgreSQL)
library(formattable)
library(knitr)
library(dplyr)
library(sf)
library(leaflet)
library("leaflet.extras")
library(htmltools)
library(tidyr)
library(stringr)
library(rgdal)
library(rpostgis)
library(stringr)
library(scales)
library(colorspace)
library(geojsonsf)
library(rmapshaper)
library(tigris)

#### Prep Steps ####

### Load GIS data ###
# get census places
## pull in 2020 CBF Places ##
  places <- places(state = 'CA', year = 2020, cb = TRUE) %>% select(-c(STATEFP, PLACEFP, PLACENS, AFFGEOID, STUSPS, STATE_NAME, LSAD, ALAND, AWATER))
  city_index2 <- city_index_ %>% select(geoid, geo_name) %>% left_join(city_index, by = "geo_name")
  city_gis <- places %>% right_join(city_index2, by = c("GEOID" = "geoid"))
  
# transform for leaflet projection
city_gis <- st_transform(city_gis, CRS("+proj=longlat +datum=WGS84 +no_defs"))


### map ###
#orange for JENI 2021
#pal_impact <- #colorBin(palette = c("#800080", "orange", "red", "yellow"), domain = city_gis$quadrant, bins = 4, reverse = FALSE, pretty = FALSE)

#create popup
popup <- paste("<div class='leaflet-popup-scrolled' style='max-width:800px;max-height:200px'> City: <span style='font-size: 16px; font-weight: bold;'>", city_gis$geo_name, "</span></br>Region: ", city_gis$region, 
 "</br> Disparity Rank*: <span style='font-size: 15px; font-weight: bold;'>", city_gis$disparity_rank," | </span><span style='font-size: 14px; font-weight: bold;'>",
 city_gis$disparity_z, " (z-score)</span></br>",
 "</br> Outcome Rank*: <span style='font-size: 15px; font-weight: bold;'>", city_gis$performance_rank," | </span><span style='font-size: 14px; font-weight: bold;'>",
 city_gis$performance_z, " (z-score)</span></br>",
 "Region: ", city_gis$sd_number, " | ", city_gis$region, "</br>",
"Population: ", city_gis$total_pop, "</br>",
"</br>",
"<span style='font-size: 10px; font-style: italic;'>*Z-scores range from 0-1. A higher disparity rank means a higher level of disparity. A higher outcome rank means better overall outcomes.</span>","</br>","</div>")

# build map
leaflet(city_gis, width = "100%", height = "600px") %>%

# add base map
addProviderTiles("CartoDB.PositronNoLabels")%>%
addProviderTiles("CartoDB.PositronOnlyLabels", options = providerTileOptions(pane = "markerPane")) %>%

# add map panes
addMapPane("index_pane", zIndex = 400) %>%
addMapPane("jesi_2021", zIndex = 415) %>%

# set view and layer control
setView(-118.6368454, 34.2324274, zoom = 8.5) %>%

# # add boundaries for SPAs and SDs
# addPolygons(data = sds, fillOpacity=0, color = "black", weight = 2.2, label=~distname, group = "Supervisorial Districts", options = pathOptions(pane = "sds_pane", interactive = FALSE), highlight = highlightOptions(color = "white", weight = 3, bringToFront = TRUE)) %>%
#   addPolygons(data = spas, fillOpacity=0, color = "black", weight = 2.2, label=~spa_2012, group = "Service Planning Areas", options = pathOptions(pane = "spas_pane", interactive = FALSE), highlight = highlightOptions(color = "white", weight = 3, bringToFront = TRUE))%>%


   #  #add combined legend
   # addLegend(position = "bottomleft",
   #          title="JENI Need Level",
   #          labels = c("Lowest (0-19th Percentile)", "Low (20-39th Percentile)",
   #                     "Moderate (40-59th Percentile)", "High (60-79th Percentile)","Highest (80-100th Percentile)", "JENI Not Available"),
   #          colors= c( "#ffff00","#f2bf00", "#e68000", "#d94000","#cc0000", "#808080"), opacity=.9)%>%

  # layers control
  
    htmlwidgets::onRender("function(el, x) {
        L.control.zoom({ position: 'topright' }).addTo(this)
    }")

  hc_add_theme(city_, rc_theme) 



``` -->


