---
title: "explore_chart_options"
output: 
  html_document:
    self_contained: true
---
<style>
.html-widget {
    width: 100vw !important;
    height: 100vh !important;
}
body { margin: 0; }
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# load libraries and test data
source("W:\\RDA Team\\R\\credentials_source.R")
library(dplyr)
library(purrr)
library(highcharter)
library(htmltools)
con <- connect_to_db("mosaic")
con_rc <- connect_to_db("racecounts")

low_birthweight_asian<- dbGetQuery(con, "SELECT * FROM v7.asian_hlth_low_birthweight_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
low_birthweight_nhpi <- dbGetQuery(con, "SELECT * FROM v7.nhpi_hlth_low_birthweight_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
low_birthweight_rc <- dbGetQuery(con_rc, "SELECT * FROM v7.arei_hlth_low_birthweight_state_2025;") 

low_birthweight_asbest <- low_birthweight_rc %>% pull(asbest)

low_birthweight_rc <- low_birthweight_rc %>%
  select(ends_with("_raw"), ends_with("_rate"))

homeownership_asian <- dbGetQuery(con, "SELECT * FROM v7.asian_hous_homeownership_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate")) 
homeownership_nhpi <- dbGetQuery(con, "SELECT * FROM v7.nhpi_hous_homeownership_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
homeownership_rc <- dbGetQuery(con_rc, "SELECT * FROM v7.arei_hous_homeownership_state_2025;") 

homeownership_asbest <- homeownership_rc %>% pull(asbest)
homeownership_rc <- homeownership_rc %>%
  select(ends_with("_raw"), ends_with("_rate"))

dbDisconnect(con)
dbDisconnect(con_rc)

```

```{r}
# get data in right shape
# should have at minimum: group (asian, nhpi, total); subgroup (bangladesh, philippines, etc); count, rate

# colnames(low_birthweight_asian)
# colnames(homeownership_asian)

low_bw_asian_reshape <- low_birthweight_asian %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "asian") %>%
  select(group, subgroup, count, rate) 

low_bw_nhpi_reshape <- low_birthweight_nhpi %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "nhpi") %>%
  select(group, subgroup, count, rate)

low_bw_rc_reshape <- low_birthweight_rc %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "total") %>%
  select(group, subgroup, count, rate) 

low_birthweight <- rbind(low_bw_asian_reshape, low_bw_rc_reshape) # low_bw_nhpi_reshape, 

# homeownership_asian_reshape <- homeownership_asian %>%
#    pivot_longer(
#     cols = matches("_raw$|_rate$"),
#     names_to = c("subgroup", ".value"),
#     names_pattern = "^(.+)_(raw|rate)$") %>%
#   rename(count = raw) %>%
#   mutate(group = "asian") %>%
#   select(group, subgroup, count, rate) 
# 
# homeownership_nhpi_reshape <- homeownership_nhpi %>%
#    pivot_longer(
#     cols = matches("_raw$|_rate$"),
#     names_to = c("subgroup", ".value"),
#     names_pattern = "^(.+)_(raw|rate)$") %>%
#   rename(count = raw) %>%
#   mutate(group = "nhpi") %>%
#   select(group, subgroup, count, rate) 
# 
# homeownership_rc_reshape <- homeownership_rc %>%
#   pivot_longer(
#     cols = matches("_raw$|_rate$"),
#     names_to = c("subgroup", ".value"),
#     names_pattern = "^(.+)_(raw|rate)$") %>%
#   rename(count = raw) %>%
#   mutate(group = "total") %>%
#   select(group, subgroup, count, rate) 
# 
# homeownership <- rbind(homeownership_asian_reshape, homeownership_nhpi_reshape, homeownership_rc_reshape)

# Set colors for bars
base_colors <- c(
  "asian" = "#070024", 
  "total" = "#ac0068",
  "default" = "#cccccc")

drilldown_colors <- c(
  "asian subgroups" = "#7cb5ec", 
  "total" = "#ac0068",
  "default" = "#cccccc"
)

color_map <- c(
  "asian subgroups" = "#7cb5ec", 
  "asian" = "#070024",             
  "other" = "#ac0068",
  "total" = "#ac0068",
  "default" = "#cccccc"
)

```

Current Mockup: https://catalystcalifornia.sharepoint.com/:p:/s/Portal/IQB2I72xqiLbQYADd3kq9sa8AZMzOGt4QlRKnGaOnVPirrY?e=vzYzR7

This document aims to test out available chart options for disaggregation. For interactive charts, RDA primarily uses the library Highcharts which is the starting point for the working examples below. We should assess the following charts and provide feedback. Based on feedback we can assess if Highcharts is an appropriate library for our needs, or if we should consider another library, or perhaps pursue no interactivity at all.


# Working Examples

## Goal 1 Example
Make the case for disaggregation. Highlight racial disparities faced by CA’s AAPI communities that would be "hidden" without disaggregated data. 

This is an example of a "drilldown" chart. You can select Asian or Total and the chart will drilldown to show the subgroups in each. Once you drilldown, you have another option to toggle to counts or stay on rates. When drilling down and viewing rates, a vertical line will appear with the state average. 

```{r}
# A bit more indicator prep for chart mock ups
# Test highcharter

low_birthweight_chart1 <- low_birthweight %>%
  # remove nphi for now
  filter(group!="nhpi") %>%
  # mutate(subgroup = ifelse(group=="nhpi", sprintf("%s (nhpi)", subgroup), subgroup)) %>%
  # mutate(group= ifelse(group=="nhpi", "asian", group)) %>%
  # round rates
  mutate(rate=round(rate, 2))

low_birthweight <- low_birthweight %>%
  filter(group!="nhpi") %>%
  mutate(
    group = case_when(group=="total" & subgroup != "total" ~ "race counts groups",
                      group=="asian" & subgroup != "total" ~ "asian subgroups",
                      group=="asian" & subgroup=="total" ~ "asian average",
                      .default=group),
    subgroup = case_when(group=="asian average" ~ "asian average",
                         .default=subgroup)
  ) %>%
  filter(group!="total"&subgroup!="total") %>%
  # mutate(subgroup = ifelse(group=="nhpi", sprintf("%s (nhpi)", subgroup), subgroup)) %>%
  # mutate(group= ifelse(group=="nhpi", "asian", group)) %>%
  mutate(rate=round(rate, 2))

# use indicator best direction to arrange "shared struggle" groups at top
if(low_birthweight_asbest=="min") {
  low_birthweight <- low_birthweight %>%
    arrange(desc(rate))
  low_birthweight_chart1 <- low_birthweight_chart1 %>%
    arrange(desc(rate))
  
} else{
  low_birthweight <- low_birthweight %>%
    arrange(rate)
  low_birthweight_chart1 <- low_birthweight_chart1 %>%
    arrange(desc(rate))
}

```

```{r}

base_bw_df <- low_birthweight_chart1 %>%
  filter(subgroup=="total") %>%
  select(-subgroup) %>%
  mutate(color = case_when(
    group == "asian" ~ color_map[["asian"]],
    group == "total" ~ color_map[["total"]],
    TRUE ~ base_colors[["default"]]
  ))
drilldown_bw_df <- low_birthweight_chart1 %>%
  filter(subgroup!="total") %>%
  group_nest(group) %>%
  mutate(
    id = group,
    type = "bar",
    # Assign color based on group
    color = case_when(
      group == "asian" ~ color_map[["asian subgroups"]],
      group == "total" ~ color_map[["total"]],
      .default = color_map[["default"]]
    ),
    # Store both rate-sorted and count-sorted versions
    data = map(data, function(df) {
      df %>%
        arrange(desc(rate)) %>%
        mutate(name = subgroup, y = rate, y2 = count) %>%
        select(name, y, y2) %>%
        list_parse()
    }),
    data_count_sorted = map2(data, group, function(df, grp) {
      low_birthweight_chart1 %>%
        filter(group == grp, subgroup != "total") %>%
        arrange(desc(count)) %>%
        mutate(name = subgroup, y = count, y2 = rate) %>%
        select(name, y, y2) %>%
        list_parse()
    })
  ) %>%
  select(id, type, color, data, data_count_sorted)

total_rate_bw <- base_bw_df %>%
  filter(group=="total") %>%
  pull(rate) %>%
  round(2)

drilldown_bw_js <- paste0("function(e) {
  var chart = this;
  
  // Store the original drilldown data before any modifications
  var seriesId = e.seriesOptions.id;
  
  if (!chart.originalDrilldownData) {
    chart.originalDrilldownData = {};
  }
  
  // Store both rate and count sorted data from the drilldown series options
  if (e.seriesOptions.data && e.seriesOptions.data_count_sorted) {
    chart.originalDrilldownData[seriesId] = {
      rateSorted: JSON.parse(JSON.stringify(e.seriesOptions.data)),
      countSorted: JSON.parse(JSON.stringify(e.seriesOptions.data_count_sorted))
    };
  }
  
  // Add plot line (only for rates view)
  chart.yAxis[0].addPlotLine({
    id: 'total-line',
    value: ", total_rate_bw, ",
    color: '#FF0000',
    width: 2,
    dashStyle: 'Dash',
    zIndex: 5,
    label: { text: 'State Average: ", total_rate_bw, "', align: 'left', rotation: 0, y: 5, x: 5 }
  });
  
  // Reset state - always start with rates
  chart.currentView = 'rates';
  
  // Update data label formatter for rates (initial state)
  chart.update({
    plotOptions: {
      bar: {
        dataLabels: {
          formatter: function() {
            return this.y;
          }
        }
      }
    }
  }, false);
  
  chart.yAxis[0].update({title: {text: 'Rate'}}, false);
  
  // Add toggle button if it doesn't exist
  if (!chart.customToggleButton) {
    chart.customToggleButton = chart.renderer.button(
      'Show Counts',
      chart.plotLeft + 10,
      10,
      function() {
        if (chart.currentView === 'rates') {
          // Switch to counts
          chart.currentView = 'counts';
          chart.yAxis[0].update({title: {text: 'Count'}}, false);
          chart.customToggleButton.attr({text: 'Show Rates'});
          
          // Remove the state average line
          chart.yAxis[0].removePlotLine('total-line');
          
          // Update all drilled-down series to show counts
          chart.series.forEach(function(s) {
            if (s.options.id && chart.originalDrilldownData[s.options.id]) {
              s.setData(chart.originalDrilldownData[s.options.id].countSorted, false);
            }
          });
          
          // Update data label formatter for counts
          chart.update({
            plotOptions: {
              bar: {
                dataLabels: {
                  formatter: function() {
                    return this.y.toLocaleString('en-US');
                  }
                }
              }
            }
          }, false);
          
        } else {
          // Switch to rates
          chart.currentView = 'rates';
          chart.yAxis[0].update({title: {text: 'Rate'}}, false);
          chart.customToggleButton.attr({text: 'Show Counts'});
          
          // Re-add the state average line
          chart.yAxis[0].addPlotLine({
            id: 'total-line',
            value: ", total_rate_bw, ",
            color: '#f91105',
            width: 2,
            dashStyle: 'Dash',
            zIndex: 5,
            label: { text: 'State Average: ", total_rate_bw, "', align: 'left', rotation: 0, y: 5, x: 5 }
          });
          
          // Update all drilled-down series to show rates
          chart.series.forEach(function(s) {
            if (s.options.id && chart.originalDrilldownData[s.options.id]) {
              s.setData(chart.originalDrilldownData[s.options.id].rateSorted, false);
            }
          });
          
          // Update data label formatter for rates
          chart.update({
            plotOptions: {
              bar: {
                dataLabels: {
                  formatter: function() {
                    return this.y;
                  }
                }
              }
            }
          }, false);
        }
        
        chart.redraw();
      },
      {
        fill: '#2196F3',
        stroke: '#1976D2',
        'stroke-width': 1,
        r: 3,
        padding: 8
      },
      {
        fill: '#1976D2'
      },
      {
        fill: '#0D47A1'
      }
    )
    .css({
      color: 'white',
      fontWeight: 'bold',
      fontSize: '12px'
    })
    .add();
  } else {
    // Button exists, reset it
    chart.customToggleButton.attr({text: 'Show Counts'});
  }
  
  chart.redraw();
}")

drillup_bw_js <- "function() { 
  var chart = this;
  chart.yAxis[0].removePlotLine('total-line');
  
  // CRITICAL: Reset all drilldown series data back to rate-sorted BEFORE drilling up
  // This ensures the next drilldown starts with the correct data
  if (chart.originalDrilldownData && chart.options.drilldown && chart.options.drilldown.series) {
    chart.options.drilldown.series.forEach(function(drilldownSeries) {
      var seriesId = drilldownSeries.id;
      if (chart.originalDrilldownData[seriesId]) {
        drilldownSeries.data = chart.originalDrilldownData[seriesId].rateSorted;
      }
    });
  }
  
  // Remove toggle button on drillup
  if (chart.customToggleButton) {
    chart.customToggleButton.destroy();
    chart.customToggleButton = null;
  }
  
  // Clear state
  chart.currentView = null;
  
  // Reset to rates view
  chart.yAxis[0].update({title: {text: 'Rate'}}, false);
  
  // Reset data label formatter
  chart.update({
    plotOptions: {
      bar: {
        dataLabels: {
          formatter: function() {
            return this.y;
          }
        }
      }
    }
  }, false);
  
  chart.redraw();
}"

hc <- hchart(
  base_bw_df,
  "bar",
  hcaes(x = group, y = rate, drilldown = group, y2 = count, color=color),
  name = "Low Birthweight Rates",
  colorByPoint = FALSE
) %>%
  hc_chart(
    height = 400, 
    scrollablePlotArea = list(minHeight = 500, opacity = 0.9),
    events = list(
      drilldown = JS(drilldown_bw_js),
      drillup = JS(drillup_bw_js)
    )
  ) %>%
  hc_tooltip(enabled = FALSE) %>%
  hc_drilldown(
    allowPointDrilldown = TRUE,
    series = list_parse(drilldown_bw_df),  
    drillUpButton = list(
      relativeTo = "spacingBox",
      position = list(
        align = "right",
        verticalAlign = "middle"
      )
    )
  ) %>%
  hc_title(
    text = "Low Birthweight by Race/Ethnicity",
    align = "center" 
  ) %>% 
  hc_subtitle(
    text = "Click the Asian or Total bar below to disaggregate data",
    align = "center"
  ) %>%
  hc_plotOptions(
    bar = list(
      dataLabels = list(
        enabled = TRUE,
        align = "right",
        inside = FALSE,
        style = list(color = "black")
      ))) %>% 
  hc_xAxis(labels = list(
    enabled = TRUE,
    style = list(
      fontSize = "11px"))) %>%
  hc_yAxis(title = list(text = "Rate")) %>%
  hc_legend(enabled=FALSE)


# Wrap with constrained width
htmltools::tagList(
  htmltools::tags$div(
    class = "chart-wrapper",
    style = "max-width: 1200px; width: 100%; margin: 0 auto;",
    htmltools::tags$div(
      class = "chart-container",
      style = "height: auto; overflow: hidden; border: 1px solid #ccc;",
      hc
    )
  ),
  htmltools::tags$script(HTML("
    function resizeChart() {
      if (typeof Highcharts !== 'undefined' && Highcharts.charts.length > 0) {
        Highcharts.charts.forEach(function(chart) {
          if (chart) {
            var wrapper = document.querySelector('.chart-wrapper');
            if (wrapper) {
              var targetWidth = wrapper.offsetWidth - 40;
              chart.setSize(targetWidth, null, false);
            }
          }
        });
      }
    }
    
    setTimeout(resizeChart, 100);
    window.addEventListener('resize', resizeChart);
  "))
)

```
#

## Goal 2 Example
Counter the model minority myth by showing the “shared struggle”. Highlight where AAPI subgroup rates are on par with other BIPOC rates. 

Here is an example of the second chart type. The initial view shows rates for all subgroups and aggregated groups (like traditional Race Counts race/ethnic groups) for initial comparison. You can interact by toggling on/off different series or by changing the view to display counts instead of rates. 

```{r}
all_subgroups <- unique(low_birthweight$subgroup)

# Create ordered categories for RATES (descending by rate)
rate_order <- low_birthweight %>%
  arrange(desc(rate)) %>%
  pull(subgroup)

# Create ordered categories for COUNTS (descending by count)
count_order <- low_birthweight %>%
  arrange(desc(count)) %>%
  pull(subgroup)

# Get unique groups for color assignment
groups <- unique(low_birthweight$group)


# Create series data for RATES with colors
series_data_rates <- low_birthweight %>%
  group_by(group) %>%
  group_split() %>%
  map(function(df) {
    data_points <- map(1:nrow(df), function(i) {
      x_index <- which(rate_order == df$subgroup[i]) - 1
      list(
        x = unname(x_index),
        y = df$rate[i],
        name = df$subgroup[i]
      )
    })
    
    # Assign color based on group
    series_color <- if(df$group[1] == "asian subgroups") {
      color_map[["asian subgroups"]]
    } else if(df$group[1] == "asian average") {
      color_map[["asian"]]
    } else {
      color_map[["other"]]  # This catches aian, latino, black, white, nhpi, etc.
    }
    
    list(
      name = df$group[1],
      data = data_points,
      id = paste0("rates-", df$group[1]),
      color = series_color
    )
  })

# Create series data for COUNTS with same colors
series_data_counts <- low_birthweight %>%
  group_by(group) %>%
  group_split() %>%
  map(function(df) {
    data_points <- map(1:nrow(df), function(i) {
      x_index <- which(count_order == df$subgroup[i]) - 1
      list(
        x = unname(x_index),
        y = df$count[i],
        name = df$subgroup[i]
      )
    })
    
    # Assign same color based on group
    series_color <- if(df$group[1] == "asian subgroups") {
      color_map[["asian subgroups"]]
    } else if(df$group[1] == "asian average") {
      color_map[["asian"]]
    } else {
      color_map[["other"]]  # This catches aian, latino, black, white, nhpi, etc.
    }
    
    list(
      name = df$group[1],
      data = data_points,
      id = paste0("counts-", df$group[1]),
      visible = FALSE,
      color = series_color
    )
  })

hc <- highchart() %>%
  hc_chart(
    type = "bar",
    reflow = TRUE,
    events = list(
      load = htmlwidgets::JS(sprintf("
        function() {
          var chart = this;
          var rateOrder = %s;
          var countOrder = %s;
          
          // NO NEED for color mapping logic anymore - colors are already set
          
          // Create toggle button
          chart.customButton = chart.renderer.button(
            'Show Counts',
            chart.plotLeft + 10,
            10,
            function() {
              var showingRates = true;
              
              // Check if rates are currently visible
              chart.series.forEach(function(s) {
                if (s.userOptions.id && s.userOptions.id.startsWith('rates-')) {
                  showingRates = s.visible;
                }
              });
              
              // Toggle all series and update x-axis
              chart.series.forEach(function(s) {
                if (s.userOptions.id) {
                  if (showingRates) {
                    // Switch to counts
                    if (s.userOptions.id.startsWith('rates-')) {
                      s.update({showInLegend: false}, false);
                      s.hide();
                    } else if (s.userOptions.id.startsWith('counts-')) {
                      s.update({showInLegend: true}, false);
                      s.show();
                    }
                  } else {
                    // Switch to rates
                    if (s.userOptions.id.startsWith('counts-')) {
                      s.update({showInLegend: false}, false);
                      s.hide();
                    } else if (s.userOptions.id.startsWith('rates-')) {
                      s.update({showInLegend: true}, false);
                      s.show();
                    }
                  }
                }
              });
              
              // Update button text, y-axis, x-axis categories, and data label formatter
              if (showingRates) {
                chart.customButton.attr({text: 'Show Rates'});
                chart.yAxis[0].update({title: {text: 'Count'}}, false);
                chart.xAxis[0].update({categories: countOrder}, false);
                chart.yAxis[0].removePlotLine('state-average-line');
                chart.update({
                  plotOptions: {
                    bar: {
                      dataLabels: {
                        formatter: function() {
                          return this.y.toLocaleString('en-US');
                        }
                      }
                    }
                  }
                }, false);
              } else {
                chart.customButton.attr({text: 'Show Counts'});
                chart.yAxis[0].update({title: {text: 'Rate'}}, false);
                chart.xAxis[0].update({categories: rateOrder}, false);
                chart.yAxis[0].addPlotLine({
                  id: 'state-average-line',
                  value: %s,
                  color: '#f91105',
                  width: 2,
                  dashStyle: 'Dash',
                  label: {text: 'State Average: %s', align: 'left', rotation: 0, y: -8, x: 5},
                  zIndex: 5
                });
                chart.update({
                  plotOptions: {
                    bar: {
                      dataLabels: {
                        formatter: function() {
                          return this.y;
                        }
                      }
                    }
                  }
                }, false);
              }
              
              chart.redraw();
            },
            {
              fill: '#2196F3',
              stroke: '#1976D2',
              'stroke-width': 1,
              r: 3,
              padding: 8
            },
            {
              fill: '#1976D2'
            },
            {
              fill: '#0D47A1'
            }
          )
          .css({
            color: 'white',
            fontWeight: 'bold',
            fontSize: '12px'
          })
          .add();
        }
      ", jsonlite::toJSON(rate_order), jsonlite::toJSON(count_order), total_rate_bw, total_rate_bw))
    )
  ) %>%
  hc_plotOptions(
    bar = list(
      dataLabels = list(
        enabled = TRUE,
        align = "right",
        inside = FALSE,
        style = list(color = "black"),
        formatter = JS("function() { return this.y; }")
      ),
      pointWidth = 20,
      grouping = FALSE,
      pointPadding = 0,
      groupPadding = 0.1
    ),
    series = 
      list(enableMouseTracking=FALSE), # removes all highlighting from hovering over bars and legend groups
      clip = FALSE) %>%
  hc_xAxis(
    categories = rate_order,
    labels = list(
      enabled = TRUE,
      style = list(fontSize = "11px")
    ),
    gridLineWidth = 1,
    gridLineColor = '#e6e6e6'
  ) %>%
  hc_yAxis(
    title = list(text = "Rate"),
    offset = 0,
    plotLines = list(
      list(
        id = 'state-average-line',
        value = total_rate_bw,
        color = '#f91105',
        width = 2,
        dashStyle = 'Dash',
        label = list(text = paste('State Average:', total_rate_bw), align = 'left', rotation=0, y=-8, x=5),
        zIndex = 5
      )
    )
  ) %>%
  hc_legend(
    enabled = TRUE,
    verticalAlign = "top",
    margin = 20
  ) %>%
  hc_tooltip(enabled = FALSE) %>%
  hc_title(
    text = "Low Birthweight by Race/Ethnicity",
    align = "center" 
  ) %>% 
  hc_subtitle(
    text = "Toggle groups on/off by clicking the legend labels below",
    align = "center"
  )

# Add RATES series (visible by default)
for (series in series_data_rates) {
  hc <- hc %>% hc_add_series(
    name = series$name,
    data = series$data,
    id = series$id,
    color = series$color  # Use pre-assigned color
  )
}

# Add COUNTS series (hidden by default)
for (series in series_data_counts) {
  hc <- hc %>% hc_add_series(
    name = series$name,
    data = series$data,
    id = series$id,
    visible = FALSE,
    showInLegend = FALSE,
    color = series$color  # Use pre-assigned color
  )
}

# Wrap with constrained width
htmltools::tagList(
  htmltools::tags$div(
    class = "chart-wrapper",
    style = "max-width: 1200px; width: 100%; margin: 0 auto;",
    htmltools::tags$div(
      class = "chart-container",
      style = "height: auto; overflow: hidden; border: 1px solid #ccc;",
      hc
    )
  ),
  htmltools::tags$script(HTML("
    function resizeChart() {
      if (typeof Highcharts !== 'undefined' && Highcharts.charts.length > 0) {
        Highcharts.charts.forEach(function(chart) {
          if (chart) {
            var wrapper = document.querySelector('.chart-wrapper');
            if (wrapper) {
              var targetWidth = wrapper.offsetWidth - 40;
              chart.setSize(targetWidth, null, false);
            }
          }
        });
      }
    }
    
    setTimeout(resizeChart, 100);
    window.addEventListener('resize', resizeChart);
  "))
)

```


## Other notes
Questions to answer regarding feasibility:
1. Would we need to use the same chart aesthetics/settings for the main (Goal 1) and the drilldown (Goal 2) charts?
For this I will explore potential limitations but we should be able to maintain similar aesthetics/chart settings if we keep separate "main" and "drilldown" charts or if those end up being split.
2. Add an instruction line like: CLICK ASIAN BAR TO SEE DISAGGREGATION/SUBGROUPS then COLLAPSE ASIAN SUBGROUPS


Other considerations
1. Asian aggregated to include asian, nhpi, and then drilldown each
2. Total line as a vertical line
3. A version where the total bar is clickable (though theoretically this would include Asian adn NHPI so we may break the logic/intuition of drilldown - will try it but better design is probably vertical line
4. Do we keep separate "main" and "drilldown" charts? How do we incorporate other groups? A toggle? A clickable total?
5. Toggle counts and rates?

Options to consider:
Highcharter: https://www.appsilon.com/post/r-highcharts-drilldown
echarts4r: https://echarts4r.john-coene.com/