---
title: "explore_chart_options"
output: 
  html_document:
    self_contained: true
---
<style>
.html-widget {
    width: 100vw !important;
    height: 100vh !important;
}
body { margin: 0; }
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# load libraries and test data
source("W:\\RDA Team\\R\\credentials_source.R")
library(dplyr)
library(purrr)
library(highcharter)
library(htmltools)
con <- connect_to_db("mosaic")
con_rc <- connect_to_db("racecounts")

low_birthweight_asian<- dbGetQuery(con, "SELECT * FROM v7.asian_hlth_low_birthweight_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
low_birthweight_nhpi <- dbGetQuery(con, "SELECT * FROM v7.nhpi_hlth_low_birthweight_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
low_birthweight_rc <- dbGetQuery(con_rc, "SELECT * FROM v7.arei_hlth_low_birthweight_state_2025;") 

low_birthweight_asbest <- low_birthweight_rc %>% pull(asbest)

low_birthweight_rc <- low_birthweight_rc %>%
  select(ends_with("_raw"), ends_with("_rate"))

homeownership_asian <- dbGetQuery(con, "SELECT * FROM v7.asian_hous_homeownership_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate")) 
homeownership_nhpi <- dbGetQuery(con, "SELECT * FROM v7.nhpi_hous_homeownership_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
homeownership_rc <- dbGetQuery(con_rc, "SELECT * FROM v7.arei_hous_homeownership_state_2025;") 

homeownership_asbest <- homeownership_rc %>% pull(asbest)
homeownership_rc <- homeownership_rc %>%
  select(ends_with("_raw"), ends_with("_rate"))

dbDisconnect(con)
dbDisconnect(con_rc)

```

```{r}
# get data in right shape
# should have at minimum: group (asian, nhpi, total); subgroup (bangladesh, philippines, etc); count, rate

# colnames(low_birthweight_asian)
# colnames(homeownership_asian)

low_bw_asian_reshape <- low_birthweight_asian %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "asian") %>%
  select(group, subgroup, count, rate) 

low_bw_nhpi_reshape <- low_birthweight_nhpi %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "nhpi") %>%
  select(group, subgroup, count, rate)

low_bw_rc_reshape <- low_birthweight_rc %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "total") %>%
  select(group, subgroup, count, rate) 

low_birthweight <- rbind(low_bw_asian_reshape, low_bw_rc_reshape) # low_bw_nhpi_reshape, 

homeownership_asian_reshape <- homeownership_asian %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "asian") %>%
  select(group, subgroup, count, rate) 

homeownership_nhpi_reshape <- homeownership_nhpi %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "nhpi") %>%
  select(group, subgroup, count, rate) 

homeownership_rc_reshape <- homeownership_rc %>%
  pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "total") %>%
  select(group, subgroup, count, rate) 

homeownership <- rbind(homeownership_asian_reshape, homeownership_nhpi_reshape, homeownership_rc_reshape)

```

Current Mockup: https://catalystcalifornia.sharepoint.com/:p:/s/Portal/IQB2I72xqiLbQYADd3kq9sa8AZMzOGt4QlRKnGaOnVPirrY?e=vzYzR7

Goal 1: Make the case for disaggregation. Highlight racial disparities faced by CA’s AAPI
communities that would be "hidden" without disaggregated data. 

Goal 2. Counter the model minority myth by showing the “shared struggle”. Highlight where
AAPI subgroup rates are on par with other BIPOC rates. 

Questions to answer regarding feasibility:
1. Would we need to use the same chart aesthetics/settings for the main (Goal 1) and the drilldown (Goal 2) charts?
For this I will explore potential limitations but we should be able to maintain similar aesthetics/chart settings if we keep separate "main" and "drilldown" charts or if those end up being split.
2. Add an instruction line like: CLICK ASIAN BAR TO SEE DISAGGREGATION/SUBGROUPS then COLLAPSE ASIAN SUBGROUPS


Other considerations
1. Asian aggregated to include asian, nhpi, and then drilldown each
2. Total line as a vertical line
3. A version where the total bar is clickable (though theoretically this would include Asian adn NHPI so we may break the logic/intuition of drilldown - will try it but better design is probably vertical line
4. Do we keep separate "main" and "drilldown" charts? How do we incorporate other groups? A toggle? A clickable total?
5. Toggle counts and rates?

Options:
Highcharter: https://www.appsilon.com/post/r-highcharts-drilldown
echarts4r: https://echarts4r.john-coene.com/

# Working Examples

## Goal 1 Example
Make the case for disaggregation. Highlight racial disparities faced by CA’s AAPI communities that would be "hidden" without disaggregated data. 

```{r}
# Test highcharter

low_birthweight_chart1 <- low_birthweight %>%
  filter(group!="nhpi") %>%
  # mutate(subgroup = ifelse(group=="nhpi", sprintf("%s (nhpi)", subgroup), subgroup)) %>%
  # mutate(group= ifelse(group=="nhpi", "asian", group)) %>%
  mutate(rate=round(rate, 2))

low_birthweight <- low_birthweight %>%
  filter(group!="nhpi") %>%
  mutate(
    group = case_when(group=="total"&subgroup!="total" ~ subgroup,
                      .default=group)
  ) %>%
  # mutate(subgroup = ifelse(group=="nhpi", sprintf("%s (nhpi)", subgroup), subgroup)) %>%
  # mutate(group= ifelse(group=="nhpi", "asian", group)) %>%
  mutate(rate=round(rate, 2))

if(low_birthweight_asbest=="min") {
  low_birthweight <- low_birthweight %>%
    arrange(desc(rate))
  low_birthweight_chart1 <- low_birthweight_chart1 %>%
    arrange(desc(rate))
  
} else{
  low_birthweight <- low_birthweight %>%
    arrange(rate)
  low_birthweight_chart1 <- low_birthweight_chart1 %>%
    arrange(desc(rate))
}
  

base_bw_df <- low_birthweight_chart1 %>%
  filter(subgroup=="total") %>%
  select(-subgroup)

drilldown_bw_df <- low_birthweight_chart1 %>%
  filter(subgroup!="total") %>%
  group_nest(group) %>%
  mutate(
    id=group,
    type="bar",
    data=map(data, mutate, name=subgroup, y2=count, y=rate),
    data=map(data, list_parse)
  )

total_rate_bw <- base_bw_df %>%
  filter(group=="total") %>%
  pull(rate) %>%
  round(2)

drilldown_bw_js <- paste0("function(e) {
  this.yAxis[0].addPlotLine({
    id: 'total-line',
    value: ", total_rate_bw, ",
    color: '#FF0000',
    width: 2,
    dashStyle: 'Dash',
    zIndex: 5,
    label: { text: 'State Average: ", total_rate_bw, "', align: 'left', rotation: 0, y: 5, x: 5 }
  });
}")

hchart(
  base_bw_df,
  "bar",
  hcaes(x = group, y = rate, drilldown = group),
  name = "Low Birthweight Rates",
  colorByPoint = TRUE
) %>%
  hc_chart(
    events = list(
      drilldown = JS(drilldown_bw_js),
      drillup = JS("function() { this.yAxis[0].removePlotLine('total-line'); }")
    )
  ) %>%
  hc_tooltip(enabled = FALSE) %>%
  hc_drilldown(
    allowPointDrilldown = TRUE,
    series = list_parse(drilldown_bw_df),
    # USE drillUpButton INSTEAD OF breadcrumbs
    drillUpButton = list(
      relativeTo = "spacingBox",
      position = list(
        align = "right",
        verticalAlign = "middle"
        # ,
        # x = 20,
        # y = -50  can use x and y for finer positioning
      )
    )
  ) %>%
  hc_title(
    text = "Low Birthweight by Race/Ethnicity",
    align = "center" 
  ) %>% 
  hc_subtitle(
    text = "Click a bar below to disaggregate",
    align = "center"
  ) %>%
  hc_plotOptions(
    bar = list(
      dataLabels = list(
        enabled = TRUE,
        align = "right",      # Alignment relative to the data point
        inside = FALSE,      # Place outside the bar
        style = list(color = "black")
      )
    )
  ) %>% 
  hc_xAxis(labels = list(
             enabled = TRUE,
             style = list(
               fontSize = "11px"))) 

# fixes needed - each bar gets a label and the chart expands to show all or atleast 
# adds a scroll bar

```
#

## Goal 2 Example
Counter the model minority myth by showing the “shared struggle”. Highlight where AAPI subgroup rates are on par with other BIPOC rates. 
```{r}
# bw_goal2 <- low_birthweight %>%
#   mutate(group= ifelse(group=="total", subgroup, group))
# 
# # Other groups as single bars
# other_groups <- bw_goal2 %>%
#   filter(group != "asian") %>%
#   select(subgroup, rate) %>%
#   rename(name = subgroup)
# 
# # Asian subgroups, excluding aggregates
# asian_subgroups <- bw_goal2 %>%
#   filter(group == "asian", !subgroup %in% c("total", "total (nhpi)")) %>%
#   select(subgroup, rate) %>%
#   rename(name = subgroup)
# 
# # Total reference line value
# total_rate <- bw_goal2 %>% filter(group=="total" & subgroup == "total") %>% pull(rate)
# 
# chart_data <- bind_rows(
#   asian_subgroups %>% mutate(color = "#0099bf", group = "Asian Subgroup"),
#   other_groups    %>% mutate(color = "#004c5f", group = "Other Group")
# ) %>%
#   arrange(rate) %>%
#   rename(y = rate) %>%  # <-- explicit mapping
#   select(name, y, color, group)          # <-- only pass what's needed
# 
# # parsed <- list_parse(chart_data)
# # str(parsed[[1]])  # should show: name, y, color, group
# 
# # # Calculate height based on number of rows (e.g., 30px per bar)
# # chart_height <- length(unique(chart_data$name)) * 30 
# 
# highchart() %>%
#   # hc_size(height=chart_height) %>%
#   hc_chart(type = "bar", 
#            height=600,
#            scrollablePlotArea = list(
#                    minHeight = 800, # Height at which scrolling kicks in
#                    opacity = 0.9,
#                    scrollPositionY = 0)) %>%
#   # Real data series
#   hc_add_series(
#     name = "All Groups",
#     data = list_parse(chart_data),
#     colorByPoint = TRUE,
#     showInLegend = FALSE
#   ) %>%
#   # Dummy series for legend toggling
#   hc_add_series(
#     name = "Asian Subgroup",
#     color = "#0099bf",
#     data = list(),
#     showInLegend = TRUE,
#     events = list(
#       legendItemClick = JS("function() {
#         var chart = this.chart;
#         var visible = this.visible !== false;
#         chart.series[0].points.forEach(function(p) {
#           if (p.group === 'Asian Subgroup') {
#             p.setVisible(!visible, false);
#           }
#         });
#         chart.redraw();
#         this.visible = !visible;
#         return false;
#       }")
#     )
#   ) %>%
#   hc_add_series(
#     name = "Other Group",
#     color = "#004c5f",
#     data = list(),
#     showInLegend = TRUE,
#     events = list(
#       legendItemClick = JS("function() {
#         var chart = this.chart;
#         var visible = this.visible !== false;
#         chart.series[0].points.forEach(function(p) {
#           if (p.group === 'Other Group') {
#             p.setVisible(!visible, false);
#           }
#         });
#         chart.redraw();
#         this.visible = !visible;
#         return false;
#       }")
#     )
#   ) %>%
#   hc_xAxis(type = "category") %>%
#   hc_yAxis(
#     plotLines = list(list(
#       id = "total-line",
#       value = total_rate,
#       color = "#FF0000",
#       width = 2,
#       dashStyle = "Dash",
#       zIndex = 5,
#       label = list(
#         text = paste("Total:", total_rate),
#         rotation = 0,
#         y = 5,
#         x=0,
#         align = "left"
#       )
#     ))
#   )


```

Here is an example of the second chart type. It's interactive in a different way but the initial view shows rates for all subgroups and aggregated groups (where relevant) for initial comparison. You can interact by toggling on/off different series or by changing the view to display counts. 
```{r}
# trying out a bar chart (not drill down) with fixed bar widths where every bar is labeled and a scroll is added
# Create a mapping of all unique subgroups to indices
all_subgroups <- unique(low_birthweight$subgroup)

# Prepare data by group with proper x-axis mapping
series_data <- low_birthweight %>%
  group_by(group) %>%
  group_split() %>%
  map(function(df) {
    # For each row, find the index of its subgroup in all_subgroups
    data_points <- map(1:nrow(df), function(i) {
      x_index <- which(all_subgroups == df$subgroup[i]) - 1  # 0-indexed for Highcharts
      list(
        x = unname(x_index),
        y = df$rate[i],
        name = df$subgroup[i]
      )
    })
    
    list(
      name = df$group[1],
      data = data_points
    )
  })

hc <- highchart() %>%
  hc_chart(
    type = "bar",
    reflow = TRUE
  ) %>%
  hc_plotOptions(
    bar = list(
      dataLabels = list(
        enabled = TRUE,
        align = "right",      # Alignment relative to the data point
        inside = FALSE,      # Place outside the bar
        style = list(color = "black")
      ),
      pointWidth = 20,
      grouping = FALSE,  # Disable grouping so bars don't cluster
      pointPadding = 0,  # Remove padding between bars
      groupPadding = 0   # Remove padding between groups
    ),
    series = list(
      clip = FALSE,
      events = list(
        legendItemClick = htmlwidgets::JS("
          function() {
            var chart = this.chart;
            var series = this;
            
            setTimeout(function() {
              // After toggle, check which categories have visible data
              var visibleCategories = new Set();
              
              chart.series.forEach(function(s) {
                if (s.visible) {
                  s.data.forEach(function(point) {
                    visibleCategories.add(point.x);
                  });
                }
              });
              
              // Show/hide axis labels based on visible data
              chart.xAxis[0].categories.forEach(function(cat, i) {
                var tick = chart.xAxis[0].ticks[i];
                if (tick && tick.label) {
                  if (visibleCategories.has(i)) {
                    tick.label.show();
                  } else {
                    tick.label.hide();
                  }
                }
              });
            }, 10);
          }
        ")
      )
    )
  ) %>%
  hc_xAxis(
    categories = all_subgroups,
    labels = list(
      enabled = TRUE,
      style = list(
        fontSize = "11px"
      )
    )
  ) %>%
  hc_yAxis(
    title = list(text = "Rate"),
    offset = 0
  ) %>%
  hc_legend(enabled = TRUE)

# Add each series
for (series in series_data) {
  hc <- hc %>% hc_add_series(
    name = series$name,
    data = series$data
  )
}

# Wrap with constrained width
htmltools::tagList(
  htmltools::tags$div(
    class = "chart-wrapper",
    style = "max-width: 1200px; width: 100%; margin: 0 auto;",
    htmltools::tags$div(
      class = "chart-container",
      style = "height: 600px; overflow-y: auto; overflow-x: hidden; border: 1px solid #ccc;",
      hc
    )
  ),
  htmltools::tags$script(HTML("
    function resizeChart() {
      if (typeof Highcharts !== 'undefined' && Highcharts.charts.length > 0) {
        Highcharts.charts.forEach(function(chart) {
          if (chart) {
            var wrapper = document.querySelector('.chart-wrapper');
            if (wrapper) {
              var targetWidth = wrapper.offsetWidth - 40;
              chart.setSize(targetWidth, null, false);
            }
          }
        });
      }
    }
    
    setTimeout(resizeChart, 100);
    window.addEventListener('resize', resizeChart);
  "))
)
# # create the bar chart
# hc <- highchart() %>%
#   hc_chart(
#     type = "bar",
#     reflow = TRUE  # Enable auto-reflow on resize
#     # ,
#     # # set the dynamic height here
#     # height=chart_height
#     ) %>%
#   hc_plotOptions(
#     bar = list(
#       pointWidth = 20  # Fixed width in pixels
#     ),
#     series = list(
#       clip = FALSE  # This is the key - disable clipping
#     )
#   ) %>%
#   hc_xAxis(categories = low_birthweight$subgroup,
#            labels = list(
#              enabled = TRUE,
#              style = list(
#                fontSize = "11px"))) %>%
#   hc_yAxis(
#     title = list(text = "Rate"),
#     # Also try increasing the plot area
#     offset = 0) %>%
#   hc_add_series(
#     name = "Low Birthweight Rate",
#     data = low_birthweight$rate
#   ) 
# 
# # Wrap with constrained width
# htmltools::tagList(
#   htmltools::tags$div(
#     class = "chart-wrapper",
#     style = "max-width: 1200px; width: 100%; margin: 0 auto;",  # Set max-width
#     htmltools::tags$div(
#       class = "chart-container",
#       style = "height: 600px; overflow-y: auto; overflow-x: hidden; border: 1px solid #ccc;",
#       hc
#     )
#   ),
#   htmltools::tags$script(HTML("
#     // Wait for Highcharts to fully load
#     function resizeChart() {
#       if (typeof Highcharts !== 'undefined' && Highcharts.charts.length > 0) {
#         Highcharts.charts.forEach(function(chart) {
#           if (chart) {
#             // Get the chart-wrapper width (respects max-width)
#             var wrapper = document.querySelector('.chart-wrapper');
#             if (wrapper) {
#               var targetWidth = wrapper.offsetWidth - 40; // Account for padding/scrollbar
#               chart.setSize(targetWidth, null, false);
#             }
#           }
#         });
#       }
#     }
#     
#     // Run multiple times to catch widget initialization
#     setTimeout(resizeChart, 100);
#     
#     window.addEventListener('resize', resizeChart);
#   "))
# )
```
#
#
```{r}
all_subgroups <- unique(low_birthweight$subgroup)

series_data <- low_birthweight %>%
  group_by(group) %>%
  group_split() %>%
  map(function(df) {
    data_points <- map(1:nrow(df), function(i) {
      x_index <- which(all_subgroups == df$subgroup[i]) - 1
      list(
        x = unname(x_index),
        y = df$rate[i],
        name = df$subgroup[i]
      )
    })
    
    list(
      name = df$group[1],
      data = data_points
    )
  })

hc <- highchart() %>%
  hc_chart(type = "bar", reflow = TRUE) %>%
  hc_plotOptions(
    bar = list(
      dataLabels = list(
        enabled = TRUE,
        align = "right",      # Alignment relative to the data point
        inside = FALSE,      # Place outside the bar
        style = list(color = "black")
      ),
      pointWidth = 20,
      grouping = FALSE,
      pointPadding = 0,
      groupPadding = 0.1  # Small padding between different subgroups
    ),
    series = list(clip = FALSE)
  ) %>%
  hc_xAxis(
    categories = all_subgroups,
    labels = list(
      enabled = TRUE,
      style = list(fontSize = "11px")
    ),
    # Add grid lines to visually separate categories
    gridLineWidth = 1,
    gridLineColor = '#e6e6e6'
  ) %>%
  hc_yAxis(
    title = list(text = "Rate"),
    offset = 0,
    plotLines = list(
      list(
        value = total_rate_bw,  # Add your reference line here
        color = 'red',
        width = 2,
        dashStyle = 'Dash',
        label = list(text = 'State Average', align = 'left', rotation=0,y=5,x=5),
        zIndex = 5
      )
    )
  ) %>%
  hc_legend(enabled = TRUE)

for (series in series_data) {
  hc <- hc %>% hc_add_series(
    name = series$name,
    data = series$data
  )
}

# Wrap with constrained width
htmltools::tagList(
  htmltools::tags$div(
    class = "chart-wrapper",
    style = "max-width: 1200px; width: 100%; margin: 0 auto;",
    htmltools::tags$div(
      class = "chart-container",
      style = "height: 600px; overflow-y: auto; overflow-x: hidden; border: 1px solid #ccc;",
      hc
    )
  ),
  htmltools::tags$script(HTML("
    function resizeChart() {
      if (typeof Highcharts !== 'undefined' && Highcharts.charts.length > 0) {
        Highcharts.charts.forEach(function(chart) {
          if (chart) {
            var wrapper = document.querySelector('.chart-wrapper');
            if (wrapper) {
              var targetWidth = wrapper.offsetWidth - 40;
              chart.setSize(targetWidth, null, false);
            }
          }
        });
      }
    }
    
    setTimeout(resizeChart, 100);
    window.addEventListener('resize', resizeChart);
  "))
)

```
