---
title: "explore_chart_options"
output: 
  html_document:
    self_contained: true
---
<style>
.html-widget {
    width: 100vw !important;
    height: 100vh !important;
}
body { margin: 0; }
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# load libraries and test data
source("W:\\RDA Team\\R\\credentials_source.R")
library(dplyr)
library(purrr)
library(highcharter)
library(htmltools)
con <- connect_to_db("mosaic")
con_rc <- connect_to_db("racecounts")

low_birthweight_asian<- dbGetQuery(con, "SELECT * FROM v7.asian_hlth_low_birthweight_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
low_birthweight_nhpi <- dbGetQuery(con, "SELECT * FROM v7.nhpi_hlth_low_birthweight_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
low_birthweight_rc <- dbGetQuery(con_rc, "SELECT * FROM v7.arei_hlth_low_birthweight_state_2025;") 

low_birthweight_asbest <- low_birthweight_rc %>% pull(asbest)

low_birthweight_rc <- low_birthweight_rc %>%
  select(ends_with("_raw"), ends_with("_rate"))

homeownership_asian <- dbGetQuery(con, "SELECT * FROM v7.asian_hous_homeownership_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate")) 
homeownership_nhpi <- dbGetQuery(con, "SELECT * FROM v7.nhpi_hous_homeownership_state_2025;") %>%
  select(ends_with("_raw"), ends_with("_rate"))
homeownership_rc <- dbGetQuery(con_rc, "SELECT * FROM v7.arei_hous_homeownership_state_2025;") 

homeownership_asbest <- homeownership_rc %>% pull(asbest)
homeownership_rc <- homeownership_rc %>%
  select(ends_with("_raw"), ends_with("_rate"))

dbDisconnect(con)
dbDisconnect(con_rc)

```

```{r}
# get data in right shape
# should have at minimum: group (asian, nhpi, total); subgroup (bangladesh, philippines, etc); count, rate

# colnames(low_birthweight_asian)
# colnames(homeownership_asian)

low_bw_asian_reshape <- low_birthweight_asian %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "asian") %>%
  select(group, subgroup, count, rate) 

low_bw_nhpi_reshape <- low_birthweight_nhpi %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "nhpi") %>%
  select(group, subgroup, count, rate)

low_bw_rc_reshape <- low_birthweight_rc %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "total") %>%
  select(group, subgroup, count, rate) 

low_birthweight <- rbind(low_bw_asian_reshape, low_bw_nhpi_reshape, low_bw_rc_reshape)

homeownership_asian_reshape <- homeownership_asian %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "asian") %>%
  select(group, subgroup, count, rate) 

homeownership_nhpi_reshape <- homeownership_nhpi %>%
   pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "nhpi") %>%
  select(group, subgroup, count, rate) 

homeownership_rc_reshape <- homeownership_rc %>%
  pivot_longer(
    cols = matches("_raw$|_rate$"),
    names_to = c("subgroup", ".value"),
    names_pattern = "^(.+)_(raw|rate)$") %>%
  rename(count = raw) %>%
  mutate(group = "total") %>%
  select(group, subgroup, count, rate) 

homeownership <- rbind(homeownership_asian_reshape, homeownership_nhpi_reshape, homeownership_rc_reshape)

```

Current Mockup: https://catalystcalifornia.sharepoint.com/:p:/s/Portal/IQB2I72xqiLbQYADd3kq9sa8AZMzOGt4QlRKnGaOnVPirrY?e=vzYzR7

Goal 1: Make the case for disaggregation. Highlight racial disparities faced by CA’s AAPI
communities that would be "hidden" without disaggregated data. 

Goal 2. Counter the model minority myth by showing the “shared struggle”. Highlight where
AAPI subgroup rates are on par with other BIPOC rates. 

Questions to answer regarding feasibility:
1. Would we need to use the same chart aesthetics/settings for the main (Goal 1) and the drilldown (Goal 2) charts?
For this I will explore potential limitations but we should be able to maintain similar aesthetics/chart settings if we keep separate "main" and "drilldown" charts or if those end up being split.
2. Add an instruction line like: CLICK ASIAN BAR TO SEE DISAGGREGATION/SUBGROUPS then COLLAPSE ASIAN SUBGROUPS


Other considerations
1. Asian aggregated to include asian, nhpi, and then drilldown each
2. Total line as a vertical line
3. A version where the total bar is clickable (though theoretically this would include Asian adn NHPI so we may break the logic/intuition of drilldown - will try it but better design is probably vertical line
4. Do we keep separate "main" and "drilldown" charts? How do we incorporate other groups? A toggle? A clickable total?
5. Toggle counts and rates?

Options:
Highcharter: https://www.appsilon.com/post/r-highcharts-drilldown

```{r}
# Test highcharter

low_birthweight <- low_birthweight %>%
  mutate(subgroup = ifelse(group=="nhpi", sprintf("%s (nhpi)", subgroup), subgroup)) %>%
  mutate(group= ifelse(group=="nhpi", "asian", group)) %>%
  arrange(rate)
  

base_bw_df <- low_birthweight %>%
  filter(subgroup=="total") %>%
  select(-subgroup)

drilldown_bw_df <- low_birthweight %>%
  filter(subgroup!="total") %>%
  group_nest(group) %>%
  mutate(
    id=group,
    type="bar",
    data=map(data, mutate, name=subgroup, y2=count, y=rate),
    data=map(data, list_parse)
  )

total_rate_bw <- base_bw_df %>%
  filter(group=="total") %>%
  pull(rate) %>%
  round(2)

drilldown_bw_js <- paste0("function(e) {
  this.yAxis[0].addPlotLine({
    id: 'total-line',
    value: ", total_rate_bw, ",
    color: '#FF0000',
    width: 2,
    dashStyle: 'Dash',
    zIndex: 5,
    label: { text: 'Total: ", total_rate_bw, "', align: 'left', rotation: 0, y: 5, x: 0 }
  });
}")

hchart(
  base_bw_df,
  "bar",
  hcaes(x = group, y = rate, drilldown = group),
  name = "Low Birthweight Rates",
  colorByPoint = TRUE
)  |>
  hc_chart(events = list(
    drilldown = JS(drilldown_bw_js),
    drillup = JS("function() { this.yAxis[0].removePlotLine('total-line'); }")
  ))|>
  hc_drilldown(
    allowPointDrilldown = TRUE,
    series = list_parse(drilldown_bw_df)
  ) 

# fixes needed - each bar gets a label and the chart expands to show all or atleast 
# adds a scroll bar

```



```{r}
bw_goal2 <- low_birthweight %>%
  mutate(group= ifelse(group=="total", subgroup, group))

# Other groups as single bars
other_groups <- bw_goal2 |>
  filter(group != "asian") |>
  select(subgroup, rate) |>
  rename(name = subgroup)

# Asian subgroups, excluding aggregates
asian_subgroups <- bw_goal2 |>
  filter(group == "asian", !subgroup %in% c("total", "total (nhpi)")) |>
  select(subgroup, rate) |>
  rename(name = subgroup)

# Total reference line value
total_rate <- bw_goal2 |> filter(group=="total" & subgroup == "total") |> pull(rate)

chart_data <- bind_rows(
  asian_subgroups |> mutate(color = "#0099bf", group = "Asian Subgroup"),
  other_groups    |> mutate(color = "#004c5f", group = "Other Group")
) |>
  arrange(rate) |>
  rename(y = rate) |>  # <-- explicit mapping
  select(name, y, color, group)          # <-- only pass what's needed

# parsed <- list_parse(chart_data)
# str(parsed[[1]])  # should show: name, y, color, group

# # Calculate height based on number of rows (e.g., 30px per bar)
# chart_height <- length(unique(chart_data$name)) * 30 

highchart() |>
  # hc_size(height=chart_height) |>
  hc_chart(type = "bar", 
           height=600,
           scrollablePlotArea = list(
                   minHeight = 800, # Height at which scrolling kicks in
                   opacity = 0.9,
                   scrollPositionY = 0)) |>
  # Real data series
  hc_add_series(
    name = "All Groups",
    data = list_parse(chart_data),
    colorByPoint = TRUE,
    showInLegend = FALSE
  ) |>
  # Dummy series for legend toggling
  hc_add_series(
    name = "Asian Subgroup",
    color = "#0099bf",
    data = list(),
    showInLegend = TRUE,
    events = list(
      legendItemClick = JS("function() {
        var chart = this.chart;
        var visible = this.visible !== false;
        chart.series[0].points.forEach(function(p) {
          if (p.group === 'Asian Subgroup') {
            p.setVisible(!visible, false);
          }
        });
        chart.redraw();
        this.visible = !visible;
        return false;
      }")
    )
  ) |>
  hc_add_series(
    name = "Other Group",
    color = "#004c5f",
    data = list(),
    showInLegend = TRUE,
    events = list(
      legendItemClick = JS("function() {
        var chart = this.chart;
        var visible = this.visible !== false;
        chart.series[0].points.forEach(function(p) {
          if (p.group === 'Other Group') {
            p.setVisible(!visible, false);
          }
        });
        chart.redraw();
        this.visible = !visible;
        return false;
      }")
    )
  ) |>
  hc_xAxis(type = "category") |>
  hc_yAxis(
    plotLines = list(list(
      id = "total-line",
      value = total_rate,
      color = "#FF0000",
      width = 2,
      dashStyle = "Dash",
      zIndex = 5,
      label = list(
        text = paste("Total:", total_rate),
        rotation = 0,
        y = 5,
        x=0,
        align = "left"
      )
    ))
  )


```

Here is an example of the second chart type. It's interactive in a different way but the initial view shows rates for all subgroups and aggregated groups (where relevant) for initial comparison. You can interact by toggling on/off different series or by changing the view to display counts. 
```{r}
# trying out a bar chart (not drill down) with fixed bar widths where every bar is labeled and a scroll is added


# calculate a dynamic height based on number of bars
n_bars <- nrow(low_birthweight)
bar_height <- 15  # pixels per bar (adjust as needed)
chart_height <- n_bars * bar_height + 50  # Increased buffer for margins

# create the bar chart
hc <- highchart() %>%
  hc_chart(
    type = "bar",
    reflow = TRUE  # Enable auto-reflow on resize
    # ,
    # # set the dynamic height here
    # height=chart_height
    ) %>%
  hc_plotOptions(
    bar = list(
      pointWidth = 20  # Fixed width in pixels
    ),
    series = list(
      clip = FALSE  # This is the key - disable clipping
    )
  ) %>%
  hc_xAxis(categories = low_birthweight$subgroup,
           labels = list(
             enabled = TRUE,
             style = list(
               fontSize = "11px"))) %>%
  hc_yAxis(
    title = list(text = "Rate"),
    # Also try increasing the plot area
    offset = 0) %>%
  hc_add_series(
    name = "Low Birthweight Rate",
    data = low_birthweight$rate
  ) 

# Wrap with constrained width
htmltools::tagList(
  htmltools::tags$div(
    class = "chart-wrapper",
    style = "max-width: 1200px; width: 100%; margin: 0 auto;",  # Set max-width
    htmltools::tags$div(
      class = "chart-container",
      style = "height: 600px; overflow-y: auto; overflow-x: hidden; border: 1px solid #ccc;",
      hc
    )
  ),
  htmltools::tags$script(HTML("
    // Wait for Highcharts to fully load
    function resizeChart() {
      if (typeof Highcharts !== 'undefined' && Highcharts.charts.length > 0) {
        Highcharts.charts.forEach(function(chart) {
          if (chart) {
            // Get the chart-wrapper width (respects max-width)
            var wrapper = document.querySelector('.chart-wrapper');
            if (wrapper) {
              var targetWidth = wrapper.offsetWidth - 40; // Account for padding/scrollbar
              chart.setSize(targetWidth, null, false);
            }
          }
        });
      }
    }
    
    // Run multiple times to catch widget initialization
    setTimeout(resizeChart, 100);
    
    window.addEventListener('resize', resizeChart);
  "))
)
```
#
#


