---
title: ""
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

## adapted from: W:/Project/RJS/LAPD Reform/Github/EMG/lapd_stops/Charts/traffic_stops_by_race_chart.Rmd

## visualize the change in hate crime experienced over time by race
# key dates: COVID-19 lock-down starting in March 2020...

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(sf)
library(scales)
library(highcharter)
library(ggplot2)

lang <- getOption("highcharter.lang")
lang$thousandsSep <- ","
options(highcharter.lang = lang)

#Add database connection and source script with functions for pushing to postgres
source("W:\\RDA Team\\R\\credentials_source.R")
options(scipen = 999)

#### Step 1: Pull in data ####
source("./calcs.R")
df <- final_df %>%
  arrange(chrono)

# need to update code below to just rename within the chart fx rather than in the df
df <- df %>%
  # clean up labels for visuals
  mutate(race_label=case_when(
    race=="latino" ~ "Latine",
    race=="aian" ~ "AIAN",
    race=="nh_black" ~ "Black",
    race=="nhpi" ~ "PI",
    race=="swana" ~ "SWANA",
    race=="nh_white" ~ "White",
    race=="nh_asian" ~ 'Asian',
    race=="total" ~ 'Total',
      TRUE~NA)) %>% #,
    #year_label=ifelse(year_=='pre-2020', 'the years before 2020', year_)) %>%
  filter(year_!='pre-2020')


## COLORS ## source: W:\RDA Team\Communications\Catalyst California Brand Guide_September 2022.pdf
orange <- "#F25922"     # 01 row on data vis colors page & brand colors page
peridot <- "#CEEA01"    # 01 row on data vis colors page & brand colors page
lavender <- "#CEC4E2"   # brand colors page
meteorite <- "#3A207D"  # brand colors page
ccblue <- "#0860BC"     # 00 row on data vis colors page 
ccyellow <-"#FFCC00"    # 01 row on data vis colors page
ccmagenta <- "#733256"  # 01 row on data vis colors page
ccpenny <- "#FF9873"    # 03 row on data vis colors page

black <- "#000000"      # brand colors page
alabaster<-"#FBFBFB"    # brand colors page
gainsboro <- "#DEDEDE"  # brand colors page

my_colors <- c('#994E77', ccblue, peridot, orange, ccpenny, lavender, ccyellow, meteorite)


#### Add data for key dates ####
# events <- data.frame(
#   df = c("March 13, 2020"),
#   label = c("First COVID-19 Lockdown")
# )

# create a category vector
x_cats <- df |>
  dplyr::distinct(year_) |>
  dplyr::arrange(year_) |>
  dplyr::pull(year_)
#x_cats <- factor(x_cats, levels = c("pre-2020", "2020", "2021", "2022"))


# build plot lines from events and the category vector
plot_lines <- lapply(seq_len(nrow(events)), function(i) {
  list(
    value = which(x_cats == events$date_[i]) - 1,
    color = gainsboro,
    width = 4,
    zIndex = 3,  
    label = list(
      text = events$label[i],
      align="left",
      rotation = 90,
      y = 10,
       padding = 1,
      backgroundColor="white",
      borderRadius=1,
      style = list(fontWeight = "bold",
                   color=black)
    )
  )
})


```

To evaluate changes in how many Californians experienced hate crimes or incidents between 2020 and 2022, we analyzed how the number and share of California Health Interview Survey (CHIS) respondents who said they had experienced a hate crime or incident. We include the date the first COVID-19 lock down began as we know that Asians experienced a surge of hate crimes and incidents during the pandemic. 

## Experiences over time

To update: **Since 2020, *the number of people stopped for minor traffic violations* decreased across race groups. Declines appeared before the pretextual stop policy took effect in March 2022.** In fact, most race groups had a slight increase in the number of stops for minor traffic violations between Q3 2022 and Q2 2023. Latine individuals had the most visible increases in stops for minor traffic violations during that period. The number of people stopped for minor traffic violations across race groups remained relatively steady in 2024. **However, there was a sharper increase in the number of people stopped for minor traffic violations for most race groups between Q1 and Q3 2025.** 

```{r}
############# STATIC VERSION ############# 


p <- ggplot(df, aes(x = year_, y = rate, color = race_label, group = race_label)) +
  geom_line(linewidth = 0.75) +
  # stable points — circles
  geom_point(data = ~dplyr::filter(., !grepl("unstable", rate_flag, ignore.case = TRUE)),
             size = 1.5, shape = 16) +
  # unstable points — triangles
  geom_point(data = ~dplyr::filter(., grepl("unstable", rate_flag, ignore.case = TRUE)),
             size = 2, shape = 17, show.legend = FALSE) +
  # hc_colors
  scale_color_manual(values = my_colors) +
  # hc_xAxis / hc_yAxis titles
  labs(
    x = "Year",
    y = "Experienced a Hate Crime / Incident (%)",
    color = NULL,
    title = "TO UPDATE: Even before Q2 2022, there was a decline in the number of\npeople stopped for minor traffic violations across race groups.",
    caption = "Race is based on survey respondent's self-identification. AIAN=American Indian or Alaska Native, NHPI=Native Hawaiian or Pacific Islander,\nSWANA=Southwest Asian or North African. AIAN, NHPI, and SWANA data include people that are one race alone or in combination with another\nrace and Latine-inclusive. Asian, Black, and White data include people that are one race alone and Latine-exclusive. Latine data include people\nof any race. In the CHIS survey, hate crimes or incidents may include times someone experienced verbal or physical abuse, or damages to their\nproperty because of their race or ethnicity.\n\nData Source: Catalyst California's calculations of California Health Interview Survey Data, 2022\n(variables AAPI19A-D)."
  ) +
  scale_x_discrete(limits = levels(x_cats)) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title          = element_text(face = "bold", hjust = 0, size = 12),
    plot.title.position = "plot",
    plot.caption        = element_text(hjust = 0, size = 7.5, color = "gray40",
                                       lineheight = 1.3),
    plot.caption.position = "plot",
    legend.position     = "bottom",
    legend.key.width    = unit(1.5, "lines")
  ) +
   annotate("text", x = Inf, y = Inf,
           label = "▲ = unstable estimate", 
           hjust = 1.1, vjust = 1.5,
           size = 3, color = "gray40")
  
p 

ggplot2::ggsave("hate_crime_race_by_year.png",
       plot = p,
       width = 2400,
       height = 1500,
       units = "px",
       dpi = 300,
       bg = "white")

```

```{r}
############# INTERACTIVE VERSION ############# 

# chart showing people who experienced hate crimes/incidents by race group
  hchart(df, type='line', 
         hcaes(x = year_, y = rate, group = race_label),
         # tooltip
         tooltip = list(headerFormat = "", 
                                pointFormat = "In <strong>{point.year_label}, {point.rate:,.0f}% of {point.race_label}</strong> adults in California experienced a hate crime or incident.<br>That rate represents <strong>{point.raw:,.0f} {point.race_label}</strong> adults.") 
)  %>%
  # colors
  hc_colors(colors=c('#994E77',ccblue,peridot,orange,lavender,ccyellow,meteorite)) %>%
  # labels
  hc_xAxis(title = list(text = "Year"),
             # add plotlines with key events
             categories = x_cats,
            plotLines = plot_lines) %>% 
  hc_yAxis(
           title = list(text = "% of People Experiencing a Hate Crime or Incident")) %>% 
  # title
  hc_title(
    text = "TO UPDATE: Even before Q2 2022, there was a decline in the number of people stopped for minor traffic violations across race groups.",
    align = "left", 
    style = list(useHTML = TRUE, fontWeight = "bold")
  ) %>% 
  # caption
  hc_caption(
    text = "Race is based on survey respondent's self-identification. AIAN=American Indian or Alaska Native, NHPI=Native Hawaiian or Pacific Islander, SWANA=Southwest Asian or North African. AIAN, NHPI, and SWANA data include people that are one race alone or in combination with another race and Latine-inclusive. Asian, Black, and White data include people that are one race alone and Latine-exclusive. Latine data include people of any race. In the CHIS survey, hate crimes or incidents may include times someone experienced verbal or
physical abuse, or damages to their property because of their race or ethnicity.
    
Data deemed unnreliable by CHIS have been are shown as..... Catalyst California's calculations of California Health Interview Survey Data, 2022 (variables AAPI19A-D).",
    align = "left",
     useHTML=TRUE, # add this
    style = list(width = "750px", # add this (set by sourceWidth minus 50) - probably good to use for all charts
                 height = "75px")) %>% # add this (set by trial/error)
  # marker and line width
  hc_plotOptions(
    line = list(
      lineWidth = 2),
    series = list(
      marker = list(
        radius = 3 )))  %>%
    hc_exporting(
    enabled = TRUE, sourceWidth=800, sourceHeight=500, 
        filename = "catalyst_california_race_counts_by_year",
     chartOptions = list(
      chart = list(
        backgroundColor = "#FFFFFF" )))

```

