---
title: ""
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
geometry: 
  - top=1.2in
  - bottom=1in
  - left=1in
  - right=1in
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{graphicx}
  - \usepackage{multicol}
  - \usepackage{xcolor}
  - \usepackage{titling}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{\includegraphics[height=0.4in]{assm_logo.png}}
  - \fancyhead[C]{\textbf{\large Brief Series Name}}
  - \fancyhead[R]{\textcolor{gray}{\today}}
  - \fancyfoot[L]{\textcolor{gray}{Confidential}}
  - \fancyfoot[C]{\textcolor{gray}{Organization Name}}
  - \fancyfoot[R]{\textcolor{gray}{Page \thepage}}
  - \renewcommand{\headrulewidth}{0.4pt}
  - \renewcommand{\footrulewidth}{0.4pt}
  - \setlength{\headheight}{50pt}
  - \fancypagestyle{plain}{\pagestyle{fancy}}
params:
  brief_title: "Default Brief Title"
  brief_subtitle: "Subtitle Here"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

\begin{center}
\LARGE \textbf{`r params$brief_title`} \\
\large `r params$brief_subtitle`
\end{center}

\vspace{0.5in}

# Executive Summary

This is your executive summary content. Single column, full width.

# Key Findings

\begin{multicols}{2}

**Finding 1:** Your first key finding goes here with supporting details and analysis.

**Finding 2:** Second finding with more detailed explanation that will wrap across the columns as needed.

**Finding 3:** Additional findings continue here. The multicols environment automatically balances the content.

\end{multicols}

# Detailed Analysis

Back to single column for detailed analysis section.

```{r sample-chart, fig.height=3, fig.width=6, fig.align='center'}
# Simple chart example (optional)
library(ggplot2)
data <- data.frame(
  Category = c("A", "B", "C", "D"),
  Value = c(25, 35, 20, 20)
)

ggplot(data, aes(x = Category, y = Value, fill = Category)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Sample Chart", y = "Percentage") +
  theme(legend.position = "none")
```

\newpage

# Page Two Content

Second page with consistent formatting. Headers and footers automatically carry over.

More content for the second page to demonstrate the template's capabilities.